---
title: "Disability Employment"
output:
  html_notebook:
    code_folding: hide
    toc: TRUE
    toc_float: TRUE
    
---

```{css, echo = FALSE}
body {
font-family: Poppins;
}
```

**Disability Employment Awareness Data Sources**

- 2022 ACS 1-year B18120 Employment Status by Disability Status and Type (Civilian Non-institutionalized Population 18 to 64 Years)
<!-- B18121 Work Experience by Disability Status and Type -->
<!-- B18130 Age by Disability Status by Poverty Status -->

- 2022 ACS 1-year B18121 Work Experience by Disability Status and Type
- 2022 ACS 1-year C18130 Poverty Status by Disability Status and Age

*ACS PUMS Variables*

- 2019 1-year, 2020 5-year, 2021 1-year, 2022 1-year

- DIS (disability flag)
- OCCP (Occupation recode based on 2018 OCC codes)
- LUM_JOBSECTOR (Modeling Sectors)
- STANDARD_JOBSECTOR (Covered Employment Sectors)

**Information**

[Census Stats](https://www.census.gov/newsroom/stories/disability-employment-awareness-month.html)

The tables include the coefficient of variation (CV) measures which is the relative amount of sampling error that is associated with a sample estimate. See [this article](https://psrc.github.io/psrccensus/articles/calculate-reliability-moe-transformed-acs.html) for interpretation.

```{r, echo = FALSE}
library(psrccensus)
library(psrcplot)
library(tidyverse)
library(tidycensus)
library(DT)
library(openxlsx)
library(echarts4r)

source('functions.R')

outdir <- "T:\\2023December\\Christy\\disability_emp"
```

```{r B18120 Employment Status, echo = FALSE}
x <- get_acs_recs('county', table.names = 'B18120', years = '2022', acs.type = 'acs1')

x02 <- x %>% 
  mutate(variable_sort = str_extract(variable, "\\d{2}$")) %>% 
  mutate(label02 = label) %>% 
  separate(label02,
           sep = ":!!",
           into = c('total', 'in_lf', 'is_emp', 'has_dis', 'dis_type')) %>%
  mutate(dis_type = ifelse(variable_sort %in% 23:28, has_dis, dis_type)) %>%
  mutate(has_dis = ifelse(variable_sort %in% 22:29, is_emp, has_dis)) %>%
  mutate(is_emp = ifelse(variable_sort %in% 22:29, NA, is_emp)) %>% 
  mutate(in_lf = str_replace_all(in_lf, ":", ""),
         is_emp = str_replace_all(is_emp, ":", ""),
         has_dis = str_replace_all(has_dis, ":", ""))
```

```{r region}

r <- x02 %>% 
  filter(name == 'Region')
```

# Disability in Region (as a whole)

```{r region with disability, echo = FALSE}
w_dis <- c('04', '13', '22')
wo_dis <- c('11', '20', '29')

r_tot <- r %>% 
  filter(variable_sort == '01') %>% 
  select(acs_type, year, name, denom_est = estimate, denom_moe = moe)

r_dis <- r %>%
  filter(variable_sort %in% c(w_dis, wo_dis)) %>% 
  group_by(name, has_dis) %>%
  summarise(estimate = sum(estimate),
            moe = moe_sum(moe = moe, estimate = estimate)) %>%
  mutate(se = moe/1.645) %>% 
  mutate(cv = (se/estimate)) %>% 
  left_join(r_tot, by = 'name') %>% 
  mutate(share_est = estimate/denom_est,
         share_moe = moe_prop(estimate, denom_est, moe, denom_moe))

r_dis_fmt <- r_dis %>% 
  select(-se, -denom_est, -denom_moe) %>% 
  relocate(any_of(c("acs_type", "year")), .before = name)

datatable(r_dis_fmt) %>%
  formatRound(columns=c('estimate'), digits = 0) %>% 
  formatRound(columns=c('moe', 'cv'), digits = 2) %>% 
  formatPercentage(c('cv', 'share_est', 'share_moe'), 1)

```

- 9.7% of people in the Central Puget Sound Region between the ages of 18-64 report having a disability 

## Disability types

```{r region disability types, echo = FALSE}

#4, 13, 22 "With disability" total

r_dis_type_denom <- r %>% 
  filter(variable_sort == '01') %>% 
  group_by(name) %>% 
  summarise(denom_est = sum(estimate),
            denom_moe = moe_sum(moe = moe, estimate = estimate))

r_dis_type <- r %>% 
  filter(has_dis == 'With a disability') %>% 
  filter(!(variable_sort %in% c('04', '13', '22'))) %>% # 444,261 
  group_by(acs_type, year, name, dis_type) %>%
  summarise(estimate = sum(estimate),
            moe = moe_sum(moe = moe, estimate = estimate)) %>% 
  mutate(se = moe/1.645) %>% 
  mutate(cv = (se/estimate)) %>% 
  left_join(r_dis_type_denom, by = 'name') %>% 
  mutate(share_est = estimate/denom_est,
         share_moe = moe_prop(estimate, denom_est, moe, denom_moe))

r_dis_type_fmt <- r_dis_type %>% 
  select(-se, -denom_est, -denom_moe)

datatable(r_dis_type_fmt)%>%
  formatRound(columns=c('estimate'), digits = 0) %>% 
  formatRound(columns=c('moe'), digits = 2) %>% 
  formatPercentage(c('cv', 'share_est', 'share_moe'), 1)

```

* Individuals can report more than one disability. Aggregating all disability types will not equal the reported 'With a disability' totals.

- Most individuals in the region with a disability report at least a cognitive difficulty followed by independent living difficulty, and an ambulatory difficulty

```{r}
ggplot(r_dis_type_fmt, aes(dis_type, share_est, fill = dis_type)) +
  geom_col(position = 'dodge') +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  geom_linerange(aes(ymax = share_est + share_moe, ymin = share_est - share_moe),
                 position = position_dodge2(width = 0, padding = 0.5)) +
  labs(y = NULL,
       x = NULL) +
  theme(legend.title = element_blank(),
        legend.position = "none")
```

## Within the Labor Force


```{r in labor force with disability}
# in labor force, emp/ump, with a disability

r_lf_emp_dis_denom <- r %>%
  filter(variable_sort %in% c('01', '02')) %>%
  mutate(label02 = label) %>% 
  mutate(label02 = str_replace_all(label02, ".*In the labor force.*", "in_lf"),
         label02 = str_replace_all(label02, ".*Total.*", "reg_tot")) %>% 
  pivot_wider(id_cols = c('name'),
              names_from = 'label02',
              names_glue = "denom_{label02}_{.value}",
              values_from = c('estimate', 'moe'))

lf_rows <- c('04', '11', '13', '20')

r_lf_emp_dis <- r %>% 
  filter(variable_sort %in% lf_rows) %>% 
  mutate(se = moe/1.645) %>% 
  mutate(cv = (se/estimate)) %>% 
  left_join(r_lf_emp_dis_denom, by = 'name') %>% 
  mutate(share_est_reg = estimate/denom_reg_tot_estimate,
         share_moe_reg = moe_prop(estimate, denom_reg_tot_estimate, moe, denom_reg_tot_moe),
         share_est_inlf = estimate/denom_in_lf_estimate,
         share_moe_inlf = moe_prop(estimate, denom_in_lf_estimate, moe, denom_in_lf_moe))

r_lf_emp_dis_fmt <- r_lf_emp_dis %>% 
  select(acs_type, year, name, in_lf, is_emp, has_dis, estimate, moe, cv, starts_with('share'))

datatable(r_lf_emp_dis_fmt) %>%
  formatRound(columns=c('estimate'), digits = 0) %>% 
  formatRound(columns=c('moe'), digits = 2) %>% 
  formatPercentage(9:13, 1)


```

Of those in the labor force

- 6.3% (approx 139,000 persons) with a disability are employed
- Individuals with a disability and unemployed represent 0.6% (12,700 persons) of the labor force

## Not in the Labor Force
```{r not in labor force with disability}
# NOT in labor force, with a disability

r_not_lf_dis_denom <- r %>%
  filter(variable_sort %in% c('01', '21')) %>%
  mutate(label02 = label) %>% 
  mutate(label02 = str_replace_all(label02, ".*Not in labor force.*", "in_lf"),
         label02 = str_replace_all(label02, ".*Total.*", "reg_tot")) %>% 
  pivot_wider(id_cols = 'name',
              names_from = 'label02',
              names_glue = "denom_{label02}_{.value}",
              values_from = c('estimate', 'moe'))

not_lf_rows <- c('22', '29')

r_not_lf_dis <- r %>% 
  filter(variable_sort %in% not_lf_rows) %>% 
  mutate(se = moe/1.645) %>% 
  mutate(cv = (se/estimate)) %>% 
  left_join(r_not_lf_dis_denom, by = 'name') %>% 
  mutate(share_est_reg = estimate/denom_reg_tot_estimate,
         share_moe_reg = moe_prop(estimate, denom_reg_tot_estimate, moe, denom_reg_tot_moe),
         share_est_inlf = estimate/denom_in_lf_estimate,
         share_moe_inlf = moe_prop(estimate, denom_in_lf_estimate, moe, denom_in_lf_moe))

r_not_lf_dis_fmt <- r_not_lf_dis %>% 
  select(acs_type, year, name, in_lf, is_emp, has_dis, estimate, moe, cv, starts_with('share'))

datatable(r_not_lf_dis_fmt) %>%
  formatRound(columns=c('estimate'), digits = 0) %>% 
  formatRound(columns=c('moe'), digits = 2) %>% 
  formatPercentage(9:13, 1)


```

<!-- X% are not in in the labor force -->

- Of those not in the labor force, 21.8% percent report a disability 

# Work Experience by Disability Status and Type

```{r B18121 Work Experience by Disability Status and Type, echo = FALSE}
we <- get_acs_recs('county', table.names = 'B18121', years = '2022', acs.type = 'acs1')

we02 <- we %>%
  mutate(variable_sort = str_extract(variable, "\\d{2}$")) %>%
  mutate(label02 = label) %>%
  separate(label02,
           sep = ":!!",
           into = c('total', 'duration', 'has_dis', 'dis_type')) %>%
  mutate(duration = str_replace_all(duration, ":", ""),
         has_dis = str_replace_all(has_dis, ":", ""))
```

```{r}

r_we <- we02 %>% 
  filter(name == 'Region')

exc_we_rows <- c('01', '02', '11', '12', '20', '21')

r_we_denom <- r_we %>% 
  filter(variable_sort %in% '01') %>% 
  select(name, denom_est = estimate, denom_moe = moe)
```

```{r}


r_we_grp <- r_we %>% 
  filter(!(variable_sort %in% exc_we_rows)) %>% 
  mutate(is_wrk = ifelse(str_detect(duration, "Worked.*"), "Worked", "Not Worked")) %>% 
  group_by(acs_type, year, name, is_wrk, has_dis) %>% 
  summarise(estimate = sum(estimate),
            moe = moe_sum(moe = moe, estimate = estimate)) %>% 
  mutate(se = moe/1.645) %>% 
  mutate(cv = (se/estimate)) %>% 
  left_join(r_we_denom, by = 'name') %>% 
  mutate(share_est = estimate/denom_est,
         share_moe = moe_prop(estimate, denom_est, moe, denom_moe))

r_we_grp_fmt <- r_we_grp %>% 
  select(acs_type, year, name, is_wrk, has_dis, estimate, moe, cv, starts_with('share'))

datatable(r_we_grp_fmt) %>%
  formatRound(columns=c('estimate'), digits = 0) %>% 
  formatRound(columns=c('moe'), digits = 2) %>% 
  formatPercentage(c('cv', 'share_est', 'share_moe'), 1)

```
```{r}
ggplot(r_we_grp_fmt, aes(x = is_wrk, y = share_est, fill = has_dis)) +
  geom_col(position = 'dodge') +
  scale_y_continuous(labels = scales::percent_format()) +
  geom_linerange(aes(ymax = share_est + share_moe, ymin = share_est - share_moe),
                 position = position_dodge2(width = 1, padding = 0.5)) +
  labs(y = NULL,
       x = NULL) +
  theme(legend.title = element_blank())


```


```{r}

# r_we_denom <- r_we %>% 
#   filter(variable_sort %in% '01') %>% 
#   select(name, denom_est = estimate, denom_moe = moe)

r_we_dis <- r_we %>% 
  filter(!(variable_sort %in% exc_we_rows)) %>% 
  group_by(acs_type, year, name, duration, has_dis) %>% 
  summarise(estimate = sum(estimate),
            moe = moe_sum(moe = moe, estimate = estimate)) %>%
  mutate(se = moe/1.645) %>% 
  mutate(cv = (se/estimate)) %>% 
  left_join(r_we_denom, by = 'name') %>% 
  mutate(share_est = estimate/denom_est,
         share_moe = moe_prop(estimate, denom_est, moe, denom_moe))

r_we_dis_fmt <- r_we_dis %>% 
  select(acs_type, year, name, duration, has_dis, estimate, moe, cv, starts_with('share'))

datatable(r_we_dis_fmt) %>%
  formatRound(columns=c('estimate'), digits = 0) %>% 
  formatRound(columns=c('moe'), digits = 2) %>% 
  formatPercentage(c('cv', 'share_est', 'share_moe'), 1)

```

```{r}
ggplot(r_we_dis_fmt, aes(x = duration, y = share_est, fill = has_dis)) +
  geom_col(position = 'dodge') +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  geom_linerange(aes(ymax = share_est + share_moe, ymin = share_est - share_moe),
                 position = position_dodge2(width = 1, padding = 0.5)) +
  labs(y = NULL,
       x = NULL) +
  theme(legend.title = element_blank())
```

<!-- # Disability Employment by Occupation (PUMS) -->
```{r PUMS test, eval = FALSE, include = FALSE}
x <- pums_vars_2021 %>% 
  filter(str_detect(var_label, ".*[d|D]isability.*"))

y <- pums_vars_2021 %>% 
  filter(str_detect(var_label, ".*[w|W]orker.*"))

pums_varsearch(".*[d|D]isability.*")
pums_varsearch("OCCP")
pums_varsearch("LUM_JOBSECTOR")


```

```{r PUMS Occp, eval=FALSE, echo=FALSE}

# COW (class of worker)
# DIS (Disability recode)
# https://psrc.github.io/psrccensus/articles/pums-1-basics.html
# https://www.census.gov/topics/health/disability/guidance/data-collection-acs.html

dis_wrkclass_pums <- get_psrc_pums(span = 1,                     # Denoting ACS 5-year estimates; 1-year also available
                                   dyear = 2021,                                 # Last data year of span
                                   level = "p",                                  # Unit of analysis == household ("p" used for person)
                                   vars = c("OCCP",                              # Occupation recode based on 2018 OCC codes
                                            # "COW",                             # Class of Worker
                                            "DIS"                                # Disability recode
                                   ))       

```

```{r, eval = FALSE, include = FALSE}

occ_lu <- read.xlsx('occp.xlsx')

dis_wrkclass_overview <- psrc_pums_count(dis_wrkclass_pums, group_vars = c("OCCP",  # Occupation recode based on 2018 OCC codes
                                                                           # "COW", # Class of Worker
                                                                           "DIS"    # Disability recode
)) 

dis_wrkclass_overview02 <- dis_wrkclass_overview %>% 
  separate(OCCP,
           into = c("occp_abbr", "occp_desc"),
           sep = "-")

dis_wrkclass_overview03 <- dis_wrkclass_overview02 %>% 
  filter(DIS != 'Total') %>% 
  group_by(DATA_YEAR, COUNTY, occp_abbr, DIS) %>% 
  summarise(count = sum(count), count_moe = moe_sum(estimate = count, moe = count_moe)) %>% 
  mutate(se = count_moe/1.645) %>% 
  mutate(cv = (se/count))

dis_wrkclass_denom <- dis_wrkclass_overview02 %>% 
  filter(occp_abbr == 'Total') %>% 
  select(COUNTY, count_denom = count, moe_denom = count_moe)

dis_wrkclass_overview04 <- dis_wrkclass_overview03 %>% 
  left_join(dis_wrkclass_denom, by = 'COUNTY') %>% 
  group_by(DATA_YEAR, COUNTY, occp_abbr, DIS) %>% 
  summarise(count, count_moe, cv, share = count/count_denom, share_moe = moe_prop(count, count_denom, count_moe, moe_denom)) %>% 
  left_join(occ_lu, by = "occp_abbr")

dis_wrkclass_overview_fmt <- dis_wrkclass_overview04 %>% 
  select(DATA_YEAR, COUNTY, occp_abbr, definition, DIS, count, count_moe, cv, share, share_moe)



```

<!-- Workers with disabilities work in various occupations -->

<!-- - 6% of those in the region who have reported a disability did not report the sector they work in -->
<!-- - The remainder of individuals who have reported a disability reported employment in various occupations. The top 3 occupations are: -->
<!--     - Office And Administrative Support Workers (.6%) -->
<!--     - Managerial roles (.5%) -->
<!--     - Sales And Related Workers (.5%) -->

<!-- **Disability status by Occupation (with disability only)** -->
```{r, eval=FALSE}
dis_wrkclass_overview_fmt_disonly <- dis_wrkclass_overview_fmt %>% 
  filter(DIS == 'With a disability')

datatable(dis_wrkclass_overview_fmt_disonly) %>%
  formatRound(columns=c('count'), digits = 0) %>% 
  formatRound(columns=c('count_moe'), digits = 2) %>% 
  formatPercentage(8:9, 1)

```
<!-- **Detailed Disability Status by Occupation** -->
```{r, eval=FALSE}
datatable(dis_wrkclass_overview_fmt) %>%
  formatRound(columns=c('count'), digits = 0) %>% 
  formatRound(columns=c('count_moe'), digits = 2) %>% 
  formatPercentage(8:9, 1)
```

# Disability Employment by Job Sectors (PUMS) 

<!-- ## PSRC Land Use Modeling Job Sectors -->
```{r PUMS model job sectors, eval=FALSE}

dis_modsect_overview_fmt_2019 <- create_modsect_pums_table(span = 1, dyear = 2019)
dis_modsect_overview_fmt_2021 <- create_modsect_pums_table(span = 1, dyear = 2021)
dis_modsect_overview_fmt_2020 <- create_modsect_pums_table(span = 5, dyear = 2020)
dis_modsect_overview_fmt_2022 <- create_modsect_pums_table(span = 1, dyear = 2022)

dis_modsect_overview_fmt <- bind_rows(dis_modsect_overview_fmt_2019, dis_modsect_overview_fmt_2021, dis_modsect_overview_fmt_2020, dis_modsect_overview_fmt_2022)

```

```{r export data, eval = FALSE}
# export as xlsx
write.xlsx(dis_modsect_overview_fmt, file.path(outdir, "disability_landusemod_pums_2019-2022.xlsx"))

```

<!-- **Detailed Disability Status by Modeling Sectors** -->
```{r, eval=FALSE}
datatable(dis_modsect_overview_fmt) %>%
  formatRound(columns=c('count'), digits = 0) %>% 
  formatRound(columns=c('count_moe'), digits = 2) %>% 
  formatPercentage(7:9, 1)
```


```{r, eval=FALSE}
dis_modsect_overview_fmt_plot <- dis_modsect_overview_fmt %>% 
  mutate(sector = str_extract(LUM_JOBSECTOR, "(?<=-).*")) %>% 
  mutate(DATA_YEAR = factor(DATA_YEAR, levels = c('2019', '2020', '2021', '2022'))) %>% 
  filter(!is.na(LUM_JOBSECTOR) & DIS == 'With a disability') # a lot of NAs for sector, focus only on disability emp

ggplot(dis_modsect_overview_fmt_plot, aes(x = reorder(sector, -share), y = share, fill = DATA_YEAR)) +
  geom_col(position = 'dodge') +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  geom_linerange(aes(ymax = share + share_moe, ymin = share - share_moe),
                 position = position_dodge2(width = 1, padding = 0.5)) +
  labs(y = NULL,
       x = NULL,
       title = 'Workers with Disabilities by Job Sector',
       caption = 'Based on sectors used for land use modeling') +
  theme(legend.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.5))

```
<!-- Workers with Disabilities by Job Sector; the sectors listed are used for PSRC's land use modeling -->

<!-- **2019** -->

<!-- **2020** -->

<!-- **2021** -->

<!-- **2022** -->



## PSRC Covered Employment Job Sectors 
```{r PUMS Covered Employment job sectors}

dis_covemp_2019 <- create_covempsect_pums_table(span = 1, dyear = 2019)
dis_covemp_2020 <- create_covempsect_pums_table(span = 5, dyear = 2020)
dis_covemp_2021 <- create_covempsect_pums_table(span = 1, dyear = 2021)
dis_covemp_2022 <- create_covempsect_pums_table(span = 1, dyear = 2022)

dis_covemp_fmt <- bind_rows(dis_covemp_2019, dis_covemp_2020, dis_covemp_2021, dis_covemp_2022)
```

```{r export to xlsx, eval=FALSE}

write.xlsx(dis_covemp_fmt, file.path(outdir, paste0("disability_covemp_pums_2019-2022_updated_", Sys.Date(),".xlsx")))
```

```{r}
datatable(dis_covemp_fmt) %>%
  formatRound(columns=c('count'), digits = 0) %>% 
  formatRound(columns=c('count_moe'), digits = 2) %>% 
  formatPercentage(8:10, 1)
```
```{r}
dis_covemp_fmt_plot <- dis_covemp_fmt %>% 
  mutate(sector = str_extract(STANDARD_JOBSECTOR, "(?<=-).*")) %>% 
  mutate(DATA_YEAR = factor(DATA_YEAR, levels = c('2019', '2020', '2021', '2022'))) %>% 
  filter(!is.na(STANDARD_JOBSECTOR) & DIS == 'With a disability') # a lot of NAs for sector, focus only on disability emp

ggplot(dis_covemp_fmt_plot, aes(x = reorder(sector, -share), y = share, fill = DATA_YEAR)) +
  geom_col(position = 'dodge') +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  geom_linerange(aes(ymax = share + share_moe, ymin = share - share_moe),
                 position = position_dodge2(width = 1, padding = 0.5)) +
  labs(y = NULL,
       x = NULL,
       title = 'Workers with Disabilities by Covered Employment Job Sectors',
       caption = 'Source: 2020 5-year ACS PUMS, 2021 1-year ACS PUMS, 2022 1-year ACS PUMS'
  ) +
  theme(legend.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.5))

```

Workers with Disabilities by Covered Employment Sector

**2019**

**2020**


**2021**

- About 2% of the region's population report a disability while working in the Services sector.


**2022**

- Increased workers with disabilities in Retail and Public Education

# Poverty by Disability Status

```{r Poverty by Disability Status}

pov_df <- get_acs_recs(geography = 'county',
                       table.names = 'C18130',
                       years = 2022,
                       acs.type = 'acs1')

pov_df_denom <- pov_df %>% filter(variable == "C18130_001" & name == 'Region') %>% select(denom = estimate, moe_denom = moe)

p_df <- pov_df %>% 
  filter(name == 'Region') %>%  #C18130_001
  mutate(denom = pov_df_denom$denom,
         moe_denom = pov_df_denom$moe_denom) %>% 
  # mutate(denom = pov_df %>% filter(variable == "C18130_001" & name == 'Region') %>% pull("estimate"),
  #        moe_denom = pov_df %>% filter(variable == "C18130_001" & name == 'Region') %>% pull("moe")) %>%
  mutate(share = estimate/denom,
         share_moe = moe_prop(num = estimate,
                              denom = denom,
                              moe_num = moe,
                              moe_denom = moe_denom))

```

## By Age Group
```{r}
p_df %>% 
  select(-denom, -moe_denom, -se) %>% 
  datatable() %>%
  formatRound(columns=c('estimate'), digits = 0) %>% 
  formatRound(columns=c('moe', 'cv'), digits = 2) %>% 
  formatPercentage(c('cv', 'share', 'share_moe'), 1)
```

```{r export Poverty to xlsx, eval=FALSE}

write.xlsx(p_df, file.path(outdir, paste0("poverty_by_disability_status_by_age_2022_", Sys.Date(),".xlsx")))
```

```{r aggregate ages}

pov_df_all_age <- pov_df %>% 
  filter(name == 'Region' & label %in% str_subset(label, ".*poverty\\s*level$")) %>% 
  mutate(label02 = label) %>% 
  separate(label02,
           sep = ":!!",
           into = c('total', 'age', 'has_dis', 'pov_level')) %>% 
  group_by(GEOID, name, state, has_dis, pov_level) %>%
  summarise(estimate = sum(estimate),
            moe = moe_sum(moe = moe, estimate = estimate)) %>% 
  mutate(se = moe/1.645) %>% 
  mutate(cv = (se/estimate)) %>% 
  bind_cols(pov_df_denom) %>% 
  mutate(share = estimate/denom,
         share_moe = moe_prop(num = estimate,
                              denom = denom,
                              moe_num = moe,
                              moe_denom = moe_denom))

```

## All Ages

```{r}
pov_df_all_age %>% 
  select(-denom, -moe_denom, -se) %>% 
  datatable() %>%
  formatRound(columns=c('estimate'), digits = 0) %>% 
  formatRound(columns=c('moe', 'cv'), digits = 2) %>% 
  formatPercentage(c('cv', 'share', 'share_moe'), 1)
```



```{r export Poverty all ages to xlsx, eval=FALSE}

write.xlsx(pov_df_all_age, file.path(outdir, paste0("poverty_by_disability_status_all_ages_2022_", Sys.Date(),".xlsx")))

```


# Regional Median Earnings by Disability Status

```{r Median Earnings by Disability Status, echo=FALSE,eval=TRUE}

# earn_df <- get_acs_recs(geography = 'county',
#                         table.names = 'B18140',
#                         years = 2022,
#                         acs.type = 'acs1')

m_earn_pums <- get_psrc_pums(span = 1, 
                             dyear = 2022, 
                             level = "p",
                             vars = c("ESR","AGEP","DIS","PERNP"))

m_earn_pums_fil <- m_earn_pums %>% 
  filter((grepl("^(Civilian|Armed) ", as.character(ESR)) & !is.na(ESR) & between(AGEP, 18, 64)))

m_earn <- psrc_pums_median(m_earn_pums_fil, 
                           stat_var = "PERNP", 
                           group_vars = c("DIS"))
```

```{r}
datatable(m_earn) %>%
  formatCurrency(columns=c(4, 5), digits = 2)
```

# Charts
```{r 2022 Labor Force by Disability Status Data}
# 1. Employment graph
df_lf <- r_lf_emp_dis_fmt %>% 
  select(acs_type:moe)

df_lf_denom <- df_lf %>% 
  group_by(acs_type, year, name, in_lf, has_dis) %>% 
  summarise(denom = sum(estimate),
            denom_moe = moe_sum(moe, estimate))

df_lf_dis <- df_lf %>% 
  left_join(df_lf_denom, by = c("acs_type", "year", "name", "in_lf", "has_dis")) %>% 
  mutate(share = estimate/denom,
         share_moe = moe_prop(num = estimate,
                              denom = denom,
                              moe_num = moe,
                              moe_denom = denom_moe))
```

```{r, eval=FALSE}

# col <- ggplot(df_lf_dis) +
#  geom_col(aes(x = has_dis, y = share, fill = is_emp),
#           position = "dodge"
#           ) +
#   facet_wrap(vars(is_emp), nrow = 1, ncol = 2)

# pie <- static_column_chart(t = df_lf_dis %>% filter(has_dis == 'With a disability'),
#                            pos = "stack",
#                            x = "has_dis",
#                            y = "share",
#                            fill = "is_emp")

# ggplot(df_lf_dis %>% filter(has_dis == 'With a disability')) +
#  geom_col(aes(x = has_dis, y = share, fill = is_emp),
#           width = 1
#           ) +
#   scale_fill_manual(values = psrc_colors$pognbgy_5[1:2]) +
#   labs(x = NULL,
#        y = NULL,
#        fill = NULL) + 
#   theme(axis.text.x = element_blank(),
#         axis.text.y = element_blank()) +
#   theme_void() +
#   coord_polar(theta = "y")

e_common(font_family = "Poppins")
# turn off legend, add width, title
p_dis <- df_lf_dis %>% filter(has_dis == 'With a disability') |>
  e_charts(is_emp) |>
  e_pie(share,
        label = list(formatter = ("{b} \n ({d}%)"),
                     position = "outside")
        ) |>
  e_color(psrc_colors$pognbgy_5[1:2]) |>
  e_title("With a Disability", left = 'center') |>
  e_legend(show = FALSE)|>
   e_text_g(
        style = list(
            text = "2022 ACS 1-year B18120 Employment Status by Disability Status and Type \n (Civilian Non-institutionalized Population 18 to 64 Years)",
            fontSize = 10,
            opacity = .7,
            fill = "grey"
        ),
        left = 0,
        bottom = 1
    )

p_wo_dis <- df_lf_dis %>% filter(has_dis == 'No disability') |>
  e_charts(is_emp) |>
  e_pie(share,
        label = list(formatter = ("{b} \n ({d}%)"),
                     position = "outside")
        ) |>
  e_color(psrc_colors$pognbgy_5[1:2])|>
  e_title("With no Disability", left = 'center')|>
  e_legend(show = FALSE)

e_p <- e_arrange(p_dis, p_wo_dis, rows = 1, cols = 2, width = "sm") 

e_p
# htmlwidgets::saveWidget(widget = e_p, file = "dis_status_by_emp_status_in_lf_22.html")
# webshot2::webshot('dis_status_by_emp_status_in_lf.html', file = 'dis_status_by_emp_status_in_lf_22.png')
```

```{r, 2022 Labor Force by Disability Status Plot}

df_lf_dis_unemp <- df_lf_dis %>% 
  filter(is_emp == 'Unemployed')

df_lf_dis_unemp_p <- static_column_chart(
  t = df_lf_dis_unemp ,
  y = "share",
  x = "has_dis",
  fill = "has_dis",
  pos = "dodge",
  xlabel = NULL,
  ylabel = NULL, 
  est = "percent",
  title = "Percent Unemployed by Disability Status 2022",
  subtitle = "",
  source = "ACS 1-year 2022, B18120 Employment Status by Disability Status and Type \n (Civilian Non-institutionalized Population 18 to 64 Years)",
  alt = NULL,
) +
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, "cm"))

df_lf_dis_unemp_p

# jpeg(file=here::here("images/B18120_in_labor_force_22.jpeg"), height = 532, width = 800)
# print(df_lf_dis_unemp_p)
# dev.off()

```

```{r  2022 CovEmp Sectors}
# 3. Table ofwith dis in one col, no dis. of sectors by percentage employed
dis_type <- unique(dis_covemp_2022$DIS)

dcemp_22 <- dis_covemp_2022 %>% 
  group_by(DIS) %>% 
  summarise(denom_count = sum(count))

dcemp_22_2 <- dis_covemp_2022 %>% 
  left_join(dcemp_22, by = 'DIS') %>% 
  select(-(cv:share_moe)) %>% 
  mutate(share = count/denom_count,
         share_moe = moe_prop(num = count, 
                              denom = denom_count, 
                              moe_num = count_moe,
                              moe_denom = moe_sum(count_moe, count)))
df <- dcemp_22_2 %>% 
  mutate(sector = case_when(STANDARD_JOBSECTOR == "1 - Const/Res" ~ "Construction & Resources",
                            STANDARD_JOBSECTOR == "2 - FIRE" ~ "Finance, Insurance \n and Real Estate (FIRE)",
                            STANDARD_JOBSECTOR == "3 - Manufacturing" ~ "Manufacturing",
                            STANDARD_JOBSECTOR == "4 - Retail" ~ "Retail",
                            STANDARD_JOBSECTOR == "5 - Services" ~ "Services",
                            STANDARD_JOBSECTOR == "6 - WTU" ~ "Wholesale Trade, Transportation, \n and Utilities",
                            STANDARD_JOBSECTOR == "7 - Government" ~ "Government",
                            STANDARD_JOBSECTOR == "8 - Public Education" ~ "Public Education")) %>% 
  mutate(sector = factor(sector, levels = c("Construction & Resources",
                                            "Finance, Insurance \n and Real Estate (FIRE)",
                                            "Manufacturing",
                                            "Retail",
                                            "Services",
                                            "Wholesale Trade, Transportation, \n and Utilities",
                                            "Government",
                                            "Public Education"))) %>% 
  mutate(DIS = ifelse(DIS == 'Without a disability', 'No disability', 'With a disability'))

# write.xlsx(dcemp_22_2, "cov_emp_22_chart.xlsx")
# df <- read.xlsx("cov_emp_22_chart.xlsx")

# equity chart facet psrcplot
df_bar <- static_bar_chart(
  t = df ,
  x = "share",
  y = "sector",
  fill = "DIS",
  pos = "dodge",
  xlabel = NULL,
  ylabel = NULL, 
  est = "percent",
  title = "Disability Status by Covered Employment Job Sectors in 2022",
  subtitle = NULL,
  source = "ACS 1-year 2022 PUMS",
  alt = NULL,
) +
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, "cm"))

df_bar

# copy and paste into console:
# jpeg(file=here::here("images/dis_status_by_covemp_22_2.jpeg"), height = 532, width = 800)
# print(df_bar)
# dev.off()
```

```{r}
# 4. Poverty/earnings

# poverty
p_df_aa <- pov_df_all_age %>% 
  select(GEOID:moe)

pov_aa_denom <- p_df_aa %>% 
  group_by(GEOID, name, state, has_dis) %>% 
  summarise(denom = sum(estimate),
            denom_moe = moe_sum(moe, estimate))

pov_df_aa <- p_df_aa %>% 
  left_join(pov_aa_denom, by = c('GEOID', 'name', 'state', 'has_dis')) %>% 
  mutate(share = estimate/denom,
         share_moe = moe_prop(num = estimate, 
                              denom = denom, 
                              moe_num = moe,
                              moe_denom = denom_moe)) %>% 
  mutate(label = case_when(pov_level == 'Income in the past 12-months at or above poverty level' ~ 'Above Poverty Level',
                           pov_level == 'Income in the past 12-months below poverty level' ~ 'Below Poverty Level'))

pdf <- pov_df_aa %>% filter(label == 'Below Poverty Level') %>% arrange(-share)

pov_df_aa_p <- static_column_chart(
  t = pdf,
  y = "share",
  x = "has_dis",
  fill = "has_dis",
  pos = "dodge",
  xlabel = NULL,
  ylabel = NULL, 
  est = "percent",
  title = "Disability Status Below Poverty Level in 2022",
  subtitle = "Income in the past 12 months",
  source = "ACS 1-year 2022, C18130 Poverty Status by Disability Status and Age",
  alt = NULL,
) +
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, "cm"))

pov_df_aa_p
# jpeg(file=here::here("images/dis_status_by_pov_lev.jpeg"), height = 532, width = 800)
# print(pov_df_aa_p)
# dev.off()  
```



