---
title: "explore_womens_history"
author: "suzanne"
date: "2023-01-19"
output: html_document
---
http://libraryarchives.metro.net/DB_Attachments/2019-0294/UnderstandingHowWomenTravel_FullReport_FINAL.pdf

distribution of trips per person per day
median
many

trip purpose

trip length

vehicle ownership

mode choice
trend
time of day

Travel Behavior Trends
Through the analysis in this report, key trends
emerge that differentiate women’s travel patterns
from men’s travel patterns, across all modes. » Across all modes, more women are making many
trips (7 or more) per day than men and more
women than men are not making any trips per day.
This means women may experience more exposure
to travel burdens (cost, stress, or safety risks), or
may be more likely to be isolated or disconnected
from the opportunities that travel affords.
» Women in Los Angeles also make shorter trips
than men, which is potentially driven by workforce
participation rates, location of employment
opportunities, and taking household-serving trips
that tend to be more localized.
» Women’s trips are more varied to a broader spread
of destinations, and are more likely to primarily
serve the needs of someone else.
» Women are more likely to live in a car-free or carlight household, take more trips with other people,
and take fewer single-occupant car trips than men.
» Women are also more likely to carpool or get a ride
from a family member or friend if they don’t have
a driver’s license.
These findings show that women may need to adjust
their own schedule and travel needs to accommodate
others, and in doing so, give up some of their own
autonomy and control over when and how they travel.
Despite these challenges and tradeoffs,
women show ingenuity in arranging their
schedules to meet their travel needs. » Women are more likely to trip-chain, or make
stops along the way to other destinations, and
describe consolidating all their errand trips into
one day where they will have access to a vehicle.
» Women in Los Angeles are also more likely than
men to travel mid-day, with a travel peak around 2
PM when transit service may be reduced.

Among female riders, almost 90% ride the system
more than three days per week.
» 57% of women bring their children on transit.
» Women ride transit because they do not have a
car, because they want to avoid traffic, or because
they do not have a license. Two of these three
reasons indicate that women who ride transit
do so because they have fewer
transportation options, and may
have less access to economic
opportunities as a result.
Still, many women do use transit
to access economic opportunity.
» Over 85% of women riders use
Metro to travel to work or school,
and of those women, 32% also use
Metro to run errands or complete
recreational trips.
Among people who make household
serving trips most frequently,
these trips comprise the same
share for women whether they use
transit or not; for men, the share of
household-serving trips declines if
they are transit users. This shows
that while men are more likely to
find alternatives to using transit to
complete household-serving trips
(using a different mode or taking
fewer trips), women are less likely
to find an alternative, and instead
work to make the transit system work for their needs.
Although the rate of adoption for TNCs like Uber and
Lyft is the same for men and women, women are more
likely than men to report that their transit use has
stayed the same as they have also begun to use TNCs.
» Women are more likely than men to say they use
TNCs for trips that transit does not serve, while
men are more likely to say they use TNCs to reach a
transit stop or station. The trips that are not served
by transit may be related to time or location, as
women’s needs differ from men’s needs by both
time of day and location.
These travel behavior findings point towards many
opportunities to adjust the services provided by Metro
to better meet the travel needs expressed by those
who are using transit. Development of a Gender Action
Plan - or a tactical plan to implement policy, design,
and service changes throughout the agency - would
help to articulate the immediate opportunities and
long-term goals that would create a system that
better serves women. Adjustments to services, vehicle
design, and policy would help minimize the time,
cost, safety, and physical burdens of riding transit
for the more than half of all riders who are women.
» The findings from Understanding
How Women Travel about women’s
mode choices, how likely they are to
travel with others in their care, and
their complex trip-chaining patterns
could all inform adjustments to
Metro’s fare policy to make it more
equitable towards women and more
cost-competitive with driving and
carpooling.
»Findings about women’s trip
purposes and primary responsibility
for household errands could all
inform the way transit vehicles,
transit stations, and bus stops are
designed, so that space for traveling
with others and carrying bags and
other belongings could be better
accommodated.
»Findings about when women are
traveling and average trip lengths
could inform new service offerings
that meet a mid-day peak travel
demand and provide better direct
connections over long distances while
minimizing transfers. 


Vehicle access issues disproportionately affect women. 

Financial access also disproportionately
affects women. Low-income women, in
particular, carry a disproportionate financial
burden when it comes to travel. 


Gender influences the spatial location of
employment opportunities, shapes women’s access
to the labor market, and, accordingly, influences
women’s transportation needs. Research shows a
relationship between occupational sex segregation
and commute distance, with women in femaledominated jobs traveling shorter distances and
working closer to home than men (Gilbert, 1998;
Hanson & Pratt, 1992, 1995; Johnston-Anumonwo,
1988). Some scholars attribute this relationship
to the spatial dispersion of feminized occupations
(Gordon et al., 1989; Hanson & Johnston, 1985).
In other words, it may be easier for women to
find local jobs compared to men. Though these
sources are decades old, they form the foundation
for the study of gendered commute patterns.
So why do women work closer to home? Statistics
suggest it has a lot to do with who takes on the
primary caregiver role. As the chart shows, before the
birth of their first child, women had slightly shorter
commutes than men on average. But in the decade
after having a first child, the average commuting time
among women fell while remaining largely the same
for fathers. The presence of children—particularly
young children—increases the number of caring
trips and the need for routes to accommodate these


» A greater number of women than men make
multiple-stop trips when traveling between their
homes and workplaces. This difference between
women and men is decreasing, however, mainly
as a result of an increase in trip chaining among
men (between 1995 and 2001, the number of
stops men made while returning home from work
increased by 24%). 


» Other demographic variables interact with gender
in predicting trip chaining. For example, having a
child under age 5 increases trip chaining by 54%
for working women and 19% for working men
as shown in the adjacent chart (McGuckin et al.,
2005a). 

Caring work: shoping, escorting?


In their household-serving trips, women carry
large shopping bags or push strollers, more often
than men (McKnight, 1994). If transit vehicles do
not have dedicated space to store bags or strollers,
women may find transit travel too inconvenient
and choose another mode. 


Some transit operators in Europe (such as
Transport for London) allocate space on transit
vehicles for strollers. However, U.S. buses
typically do not have such dedicated space;
strollers may be parked in seats but only if they
are not in use (Coale, 2015).


```{r pressure, echo=FALSE}
library(psrc.travelsurvey)
library(dplyr)
library(stringr)
library(ggplot2)
library(forcats)
library(odbc)
library(DBI)
library(tidyr)
library(psrcplot)
library(srvyr)
library(survey)



db.connect <- function(adatabase) {
  elmer_connection <- dbConnect(odbc(),
                                driver = "SQL Server",
                                server = "AWS-PROD-SQL\\SOCKEYE",
                                database = adatabase,
                                trusted_connection = "yes"
  )
}

# read table
read.dt <- function(adatabase, atable) {
  elmer_connection <- db.connect(adatabase)
  dtelm <- dbReadTable(elmer_connection, SQL(atable))
  dbDisconnect(elmer_connection)
  return(dtelm)
}

# read-in variable metadata table for levels
vars_meta <- read.dt('Elmer', 'HHSurvey.variable_metadata')


mode_vars<-c('mode_1', 'mode_simple', 'travelers_total')
other_vars<-c('final_home_rgcnum', 'hhsize', 'vehicle_count',  "hhincome_broad", 'rent_own', 'res_dur', 'student', 'education',  'hhincome_detailed', "age", "age_category", 'race_category', 'race_eth_broad', 'gender', 'employment',  'lifecycle', 'mode_acc', 'dest_purpose_cat', 'origin_purpose_cat', 'final_home_is_rgc')
trip_path_dist<-'trip_path_distance'
all_vars<-c(mode_vars, other_vars, trip_path_dist)

data_17_19<- get_hhts("2017_2019", "t", vars=all_vars)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))

data_21<- get_hhts("2021", "t", vars=all_vars)%>% mutate(year=ifelse(survey=='2017_2019', '2017/2019', '2021'))

```

```{r}

data_17_19<-data_17_19%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles"))%>%
  mutate(hhsize_simple=case_when(hhsize== '4 people' ~'4 or more people',                                                                     hhsize== '5 people' ~'4 or more people',
                                 hhsize== '6 people' ~'4 or more people',
                                 hhsize== '7 people' ~'4 or more people',
                                 hhsize== '8 people' ~'4 or more people',
                                 hhsize== '12 people' ~'4 or more people',
                                 TRUE ~ hhsize))%>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(edu_simple= case_when(education=='Bachelor degree' ~ 'Bachelors or higher', 
                               education=='Graduate/Post-graduate degree' ~ 'Bachelors or higher',
                               TRUE ~ 'Less than Bachelors degree'))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                            age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(gender_grp= case_when(gender == 'Prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                               gender=='Not listed here / prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                               gender=='Non-Binary'~ 'Non-binary, another, prefer not to answer',
                               gender=='Another'~ 'Non-binary, another, prefer not to answer',
                               TRUE ~ gender))%>%mutate(work_purpose=ifelse(dest_purpose_cat=='Work', 'Work', 'Not Work'))%>%
  
  mutate(race_short= str_extract(race_eth_broad,  "^[^ ]+"))%>%
  mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop',
                                  simple_purpose=='Escort'~ 'Escort Passenger',
                                  simple_purpose=='Errand/Other'~ 'Errands/Other',
                                  simple_purpose=='Change mode'~ 'Errands/Other',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Home' ~'Errands/Other',
                                  is.na(simple_purpose) ~ 'Errands/Other',
                                  TRUE ~ simple_purpose))%>%mutate(rgc=as.factor(final_home_is_rgc))%>%
  mutate(non_motorized_mode=ifelse((mode_simple=='Walk'|mode_simple=='Bike'),'Walk/Bike', 'Not Walk/Bike'))%>%
  mutate(mode_acc_walk=ifelse(mode_acc=='Walked or jogged', 'Walked or jogged', 'Other Access Mode'))%>%
  mutate(travelers_total_grp= case_when(travelers_total<=1 ~ '1',
                                        travelers_total==2 ~ '2',
                                        travelers_total==3 ~ '3+',
                                        travelers_total==4 ~ '4+'))


data_21<-data_21%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles"))%>%
  mutate(hhsize_simple=case_when(hhsize== '4 people' ~'4 or more people',                                                                    hhsize== '5 people' ~'4 or more people',
                                 hhsize== '6 people' ~'4 or more people',
                                 hhsize== '7 people' ~'4 or more people',
                                 hhsize== '8 people' ~'4 or more people',
                                 hhsize== '12 people' ~'4 or more people',
                                 TRUE ~ hhsize)) %>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(edu_simple= case_when(education=='Bachelor degree' ~ 'Bachelors or higher', 
                               education=='Graduate/Post-graduate degree' ~ 'Bachelors or higher',
                               TRUE ~ 'Less than Bachelors degree'))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                            age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(gender_grp= case_when(gender == 'Prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                               gender=='Not listed here / prefer not to answer' ~ 'Non-binary, another, prefer not to answer',
                               gender=='Non-Binary'~ 'Non-binary, another, prefer not to answer', 
                               gender=='Another'~ 'Non-binary, another, prefer not to answer',
                               TRUE ~ gender))%>%
  mutate(work_purpose=ifelse(dest_purpose_cat=='Work', 'Work', 'Not Work'))%>%
  mutate(simple_purpose=ifelse(dest_purpose_cat=='Home', origin_purpose_cat, dest_purpose_cat))%>%
  mutate(simple_purpose=case_when(simple_purpose=='Work'~ 'Work/School',
                                  simple_purpose=='School'~ 'Work/School',
                                  simple_purpose=='Work-related'~ 'Work/School',
                                  simple_purpose=='Shop'~ 'Shop',
                                  simple_purpose=='Escort'~ 'Escort Passenger',
                                  simple_purpose=='Errand/Other'~ 'Errands/Other',
                                  simple_purpose=='Change mode'~ 'Errands/Other',
                                  simple_purpose=='Social/Recreation' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Meal' ~ 'Social/Recreation/Meal',
                                  simple_purpose=='Home' ~'Errands/Other',
                                  is.na(simple_purpose) ~ 'Errands/Other',
                                  TRUE ~ simple_purpose))%>%mutate(rgc=as.factor(final_home_is_rgc))%>%
  mutate(non_motorized_mode=ifelse((mode_simple=='Walk'|mode_simple=='Bike'),'Walk/Bike', 'Not Walk/Bike'))%>%
  mutate(mode_acc_walk=ifelse(mode_acc=='Walked or jogged', 'Walked or jogged', 'Other Access Mode'))%>%
  mutate(travelers_total_grp= case_when(travelers_total<=1 ~ '1',
                                        travelers_total==2 ~ '2',
                                        travelers_total==3 ~ '3+',
                                        travelers_total==4 ~ '4+'))

data_17_19$hhincome_100_f=factor(data_17_19$hhincome_100, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))

data_21$hhincome_100_f=factor(data_21$hhincome_100, levels=c("Prefer not to answer",  "$100,000 or more","$75,000-$99,999", "$50,000-$74,999" ,"$25,000-$49,999" , "Under $25,000"  ))


```

```{r}
person_vars=c('gender', 'vehicle_count',  "hhincome_broad",  'hhincome_detailed', "age", "age_category", 'final_home_is_rgc', 'race_category', 'race_eth_broad', 'person_id', 'numtrips')
```

```{r}
person_data_17_19<- get_hhts("2017_2019", "p", vars=person_vars)
```


```{r}
person_data_21<- get_hhts("2021", "p", vars=person_vars)
```



```{r}
person_data_17_19<-person_data_17_19%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                             age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(numtripsgrp= case_when(numtrips==0 ~ '0',
                                numtrips<=2 ~ '1-2',
                                numtrips<=4 ~ '3-4',
                                numtrips<=5 ~ '5 or more'))
 
 

person_data_21<-person_data_21%>%mutate(NoVehicles=ifelse(vehicle_count=='0 (no vehicles)', 'No Vehicles', "Has Vehicles")) %>%
  mutate(hhincome_100= case_when(hhincome_broad=='$100,000-$199,000' ~ '$100,000 or more',
                                 hhincome_broad=='$200,000 or more' ~ '$100,000 or more',
                                 TRUE~hhincome_broad))%>%
  mutate(age_grp= case_when(age=='75-84 years' ~ '75 years or older', 
                             age == '85 or years older' ~ '75 years or older',
                            TRUE ~ age))%>%
  mutate(numtripsgrp= case_when(numtrips==0 ~ '0',
                                numtrips<=2 ~ '1-2',
                                numtrips<= 4 ~ '3-4',
                                numtrips<= 5 ~ '5 or more'))
 

person_data_17_19$hhincome_100_f=factor(person_data_17_19$hhincome_100,levels=c("Under $25,000","$25,000-$49,999",   "$50,000-$74,999","$75,000-$99,999","$100,000 or more", "Prefer not to answer"))

person_data_21$hhincome_100_f=factor(person_data_21$hhincome_100,levels=c("Under $25,000","$25,000-$49,999",   "$50,000-$74,999","$75,000-$99,999","$100,000 or more", "Prefer not to answer"))


person_data_17_19$NoVehicles=factor(person_data_17_19$NoVehicles, levels= c('No Vehicles', 'Has Vehicles'))

person_data_21$NoVehicles=factor(person_data_21$NoVehicles, levels=c('No Vehicles', 'Has Vehicles'))



simple_groupings<-c('race_eth_broad'='Race/Ethnicity','hhincome_100_f'='Household Income','NoVehicles'= 'Household Vehicles', 'age_grp'= 'Age Group', 'sample_county' ='Home County', 'age_category'='Age Category', 'final_home_is_rgc'='Home in Regional Growth Center','gender'='Gender' )
```

#number of trips
not extremely different a few subtle differences
```{r}
  
    person_summs_2017_2019 <- hhts_count(person_data_17_19,
                                        group_vars=c('gender', 'numtripsgrp'),
                                        spec_wgt='hh_weight_2017_2019_adult')%>%
  filter(numtripsgrp!='Total')%>%filter(gender %in% c('Male', 'Female'))
  

  person_summs_2021 <- hhts_count(person_data_21,
                                   group_vars=c('gender', 'numtripsgrp'),
                                   spec_wgt='person_adult_weight_2021')%>%
  filter(numtripsgrp!='Total')%>%filter(gender %in% c('Male', 'Female'))
  

static_column_chart(t= person_summs_2017_2019, x='numtripsgrp', y='share',  fill='gender', moe='share_moe', est='percent')
static_column_chart(t= person_summs_2021, x='numtripsgrp', y='share',  fill='gender', moe='share_moe', est='percent')
  
  
```



COVID may have brought more pariety to trips by purpose
```{r}

  summs_2017_2019 <- hhts_count(data_17_19,
                                        group_vars=c('gender','simple_purpose'),
                                        spec_wgt='trip_adult_weight_2017_2019')%>%drop_na(c('gender', 'simple_purpose'))%>%filter(simple_purpose!='Total')%>%filter(gender %in% c('Female', 'Male') )

  summs_2021 <- hhts_count(data_21,
                                   group_vars=c('gender','simple_purpose'),
                                   spec_wgt='trip_adult_weight_2021') %>%drop_na(c('gender', 'simple_purpose'))%>%filter(simple_purpose!='Total')%>%filter(gender %in% c('Female', 'Male') )
  
 

static_column_chart(t= summs_2017_2019, x='simple_purpose', y='share',  fill='gender', moe='share_moe', est='percent')
static_column_chart(t= summs_2021, x='simple_purpose', y='share',  fill='gender', moe='share_moe', est='percent')

```
mode shares are somewhat similar



mode
show this as % of trips without 1 person trips
```{r}

data_17_19<- data_17_19%>%drop_na(c('gender', 'travelers_total_grp'))
data_21<-data_21%>%drop_na(c('gender', 'travelers_total_grp'))

  summs_2017_2019 <- hhts_count(data_17_19,
                                        group_vars=c('gender','travelers_total_grp'),
                                        spec_wgt='trip_adult_weight_2017_2019')%>%drop_na(c('gender', 'travelers_total_grp'))%>%filter(travelers_total_grp!='Total')%>%filter(travelers_total_grp!='1')%>%filter(gender %in% c('Female', 'Male') )

  summs_2021 <- hhts_count(data_21,
                                   group_vars=c('gender','travelers_total_grp'),
                                   spec_wgt='trip_adult_weight_2021') %>%drop_na(c('gender', 'travelers_total_grp'))%>%filter(travelers_total_grp!='Total')%>%filter(travelers_total_grp!='1')%>%filter(gender %in% c('Female', 'Male') )
  
 

static_column_chart(t= summs_2017_2019, x='travelers_total_grp', y='share',  fill='gender', moe='share_moe', est='percent')
static_column_chart(t= summs_2021, x='travelers_total_grp', y='share',  fill='gender', moe='share_moe', est='percent')
```
```{r}

  summs_2017_2019 <- hhts_count(data_17_19,
                                        group_vars=c('gender','mode_simple'),
                                        spec_wgt='trip_adult_weight_2017_2019')%>%drop_na(c('gender', 'mode_simple'))%>%filter(mode_simple=='Bike')%>%filter(gender %in% c('Female', 'Male') )

  summs_2021 <- hhts_count(data_21,
                                   group_vars=c('gender','mode_simple'),
                                   spec_wgt='trip_adult_weight_2021') %>%drop_na(c('gender', 'mode_simple'))%>%filter(mode_simple=='Bike')%>%filter(gender %in% c('Female', 'Male') )
  
 

static_column_chart(t= summs_2017_2019, x='mode_simple', y='count',  fill='gender', moe='count_moe', est='number')
static_column_chart(t= summs_2021, x='mode_simple', y='count',  fill='gender', moe='count_moe', est='number')

```
