---
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: lualatex
    includes:
      in_header: trend_format_equity.tex
      extra_dependencies:
      - subfig
      - setspace
fontsize: 12pt
---

# Women's Diverse Travel Needs Often Go Overlooked
\onehalfspacing


```{r setup, include=FALSE}
### Making the Charts and pdf documents now
#install.packages("tinytex")
#update.packages(ask = FALSE, checkBuilt = TRUE)


library(rlang)
library(dplyr)
library(stringr)
library(ggplot2)
library(forcats)
library(odbc)
library(DBI)
library(tidyr)
options(tinytex.verbose = TRUE)
tinytex::tlmgr_update()
library(psrcplot)
library(tidyverse)
library(knitr)
library(magick)
library(openxlsx)
library(htmlwidgets)
library(htmltools)
library(magrittr)


source(here::here('C:/GitHub/equity-focused-trends/women_history/2023/womens_history_story_draft_datacrunching.R'))
```






```{r trip by gender, fig.dim=c(6,3), echo = FALSE}
plot<-interactive_column_chart(t= summs_2017_2019, x='simple_purpose', y='share',  fill='gender',  est='percent', title= 'Percent of Trips by  Purpose by Gender', source= 'Source: 2017/2019 PSRC Household Travel Survey',
                    color="psrc_pairs")

htmltools::save_html(plot, file="trips_by_purpose_gender.html")
```



```{r Trips by Number of Travellers, fig.dim=c(6,3), echo = FALSE}
plot<-interactive_column_chart(t= trav_summs_2017_2019, 
                    x='travelers_total_grp', 
                    y='share',  
                    fill='gender', 
                    color = 'psrc_pairs',
                    est='percent', 
                    title = 'Percent of Trips Taken by Number of Travelers, \n for People in Households with more than 2 people',  
                    source= 'Source: 2017/2019 PSRC Household Travel Survey' )

htmltools::save_html(plot, file="trips_travellers.html")
```



```{r income race and gender, fig.dim=c(8,6), echo = FALSE}
#I was having trouble getting the pums api to work, so I just took some previously downloaded data.
inc_sex_2021<-read.csv('earnings_race_gender.csv')
inc_sex_2021<- inc_sex_2021%>% mutate(PINCP_median = round(PINCP_median, -3))%>%mutate(PRACE = str_wrap(PRACE, 25))%>%
  mutate(PRACE= str_remove(PRACE, "alone"))%>%
  mutate(PRACE= str_remove(PRACE, "Alone"))

reference_val <- inc_sex_2021%>%  filter(SEX=="Total")
ref_val<- reference_val$PINCP_median

inc_sex_2021_ref<-inc_sex_2021%>%  
  filter(SEX!="Total")%>% 
  mutate(ref_val=ref_val)





```

```{r, echo=FALSE}
plot <- interactive_bar_chart(t=inc_sex_2021_ref, 
                                    x = "PINCP_median",
                                    y = "PRACE",
                                    fill = "SEX",
                                    color="psrc_pairs",
                                    est ="currency",
                                    title = "Earnings by Gender and by Race",
                                    source= '2021 PUMS earnings by race, ethnicity, and gender, for workers working more than 35 hours per week')

htmltools::save_html(plot, file="earnings_gender_race.html")
```


```{r, echo=FALSE}
plot<-interactive_column_chart(t=trips_by_mode_17_19_transit, x= 'gender', y='share', fill= 'race_eth_poc_update', title = 'Transit Mode Share by Race and Gender', source= '2017/2019 PSRC Household Travel Survey')



htmltools::save_html(plot, file="transit_gender_race.html")
```


## Women bike far less than men

\begin{flushleft}
Women bike much less than men, regardless of race and ethnicity or income. Some of this is undoubtedly because of bike network design does not feel well-suited or safe for all people of different ages and abilities. There are many considerations as to why women do not feel comfortable riding bicycles as a form of transit, as seen in a study performed by the \href{https://www.itdp.org/2022/07/06/cyclings-gender-gap/}{\underline{\textcolor{blue}{Institute for Transportation and Development Policy}}} in 2022. The inclusion of designated bike lanes, not only that increase comfort and safety for individuals, but also make traveling with children or larger cargo bikes easier would help create more accessibility and user-friendly options. Additionally, increased education about bike routes and access to bikes as a means to developing a biker's comfort and familiarity with bike safety, would help aid in bike commuting ease. 

As a result of some of these considerations not being implemented, in 2019 only \textbf{6,000} bike trips were made by non-binary folks in the central Puget Sound region, and \textbf{30,000} bike trips were made by women, but \textbf{74,000} were made by men (Figure 5). 
\end{flushleft}

```{r Bike trips, fig.dim=c(6,4), echo = FALSE}
plot<-interactive_column_chart(t= mode_summs_2017_2019, x='mode_simple', y='count',  fill='gender_grp', color = "psrc_pairs", est='number', title= "Number of Bike Trips by Gender", source = 'Source: 2019 PSRC Household Travel Survey Data')

htmltools::save_html(plot, file="bike_gender.html")
```


```{r Age group by gender, fig.dim=c(6,3), echo = FALSE}
pums19_sex_age<-read.csv('older_gender.csv')
pums19_sex_age<-pums19_sex_age%>%mutate(count=round(count, -3))%>%mutate(BIN_AGE = factor(BIN_AGE, levels = c('between 65 and 75 years', 'between 75 and 85 years', '85 years and over')))
plot<-interactive_column_chart(pums19_sex_age, 
                    x= 'BIN_AGE', 
                    y= 'count', 
                    fill= 'SEX',
                    est= 'number',
                    color = 'psrc_pairs',
                    title = 'Population by Age Group by Gender', 
                    source='Source: 2019 Public Use Microsample Census data')

htmltools::save_html(plot, file="older_gender.html")

```



