---
title: "March Women's History Month Blog Post"
date: today
format: 
  html:
    mainfont: Poppins
    theme: [cosmo, ../custom.scss]
    toc: true
    df-print: kable
    fig-asp: .7
    fig-width: 4
    warning: false
    echo: false
---

```{r}
#| include: false

library(tidyverse)
library(psrc.travelsurvey)
library(psrccensus)
library(psrcplot)
install_psrc_fonts()

source("load_data.R")

pums_source <- "Source: 2019-2023 American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS) data"
hts_source <- "Puget Sound Regional Council (PSRC) 2017-2023 Household Travel Survey"
```


# median person's earnings by gender and race and ethnicity

- only include full-time workers working more than 35hr/wk
- the results show that median earnings of male full-time workers are higher than female full-time workers in some racial groups, including 
  - Asian
  - Hispanic or Latino
  - Some Other Race
  - Two or More Races
  - White 
- while median earnings are statistically equal in other racial groups
  - American Indian or Alaskan Native
  - Black or African American
  - Native Hawaiian and Other Pacific Islander

::: {.callout-note}
categories that female and male full-time workers have statistically equal median earnings (moe bars overlapping) due to low sample size: 

- American Indian or Alaskan Native
- Black or African American
- Native Hawaiian and Other Pacific Islander
:::


:::{.panel-tabset}

## chart

```{r}
#| fig-width: 5
#| fig-asp: 0.6
#| out-width: "75%"
#| fig-align: "center"

ggplot(df_income_sex_race, aes(x=fct_rev(RAC1P_psrc), y=PERNP_35hr_median, fill=SEX_psrc)) +
  geom_col(position = position_dodge(), width = 0.8) +
  geom_text(aes(label=PERNP_35hr_format),
            hjust = -0.1,
            position = position_dodge(0.8)) +
  coord_flip() +
  scale_fill_manual(values = psrc_colors$pgnobgy_5) +
  scale_y_continuous(limits = c(0,127000)) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank())+
  labs(caption = pums_source)

```

## chart with margin of error

```{r}
#| fig-width: 5
#| fig-asp: 0.6
#| out-width: "75%"
#| fig-align: "center"

ggplot(df_income_sex_race, aes(x=fct_rev(RAC1P_psrc), y=PERNP_35hr_median, fill=SEX_psrc)) +
  geom_col(position = position_dodge(), width = 0.8) +
  geom_errorbar(aes(ymin = PERNP_35hr_median - PERNP_35hr_median_moe, ymax = PERNP_35hr_median + PERNP_35hr_median_moe),
                position = position_dodge(0.8),
                width = 0.4) +
  coord_flip() +
  scale_fill_manual(values = psrc_colors$pgnobgy_5) +
  scale_y_continuous(limits = c(0,127000)) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank())+
  labs(caption = pums_source)

```

## data

```{r}
#| tbl-cap: "median earnings by gender and race and ethnicity (workers working more than 35hr/wk) <br> [Source: 2019-2023 American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS) data]"

df_income_sex_race %>% 
  mutate(PERNP_35hr_median = round(PERNP_35hr_median)) %>%
  pivot_wider(id_cols = "RAC1P_psrc", names_from = "SEX_psrc", values_from = c("PERNP_35hr_median"))
```

:::

# share of women in occupations

- only include full-time workers working more than 35hr/wk
- female workers hold 41% of the total full-time workforce
- note that there may be a range of earnings within the same occupation. female workers in the highest paying occupation may not earn as much as male workers in the same occupation

:::{.panel-tabset}

## chart

- two versions of the same chart: one ordered by highest paying occupations, the other ordered by share of female

```{r}
#| fig-width: 4
#| fig-asp: 1
#| out-width: "100%"
#| fig-align: "center"
#| layout-ncol: 2
#| fig-cap: 
#|   - "order by highest paying occupations"
#|   - "order by share of female"

df <- df_share_workers_35hr_sex %>%
  left_join(df_income_occupation %>% 
              select(SOCP2,PERNP_35hr_median,PERNP_35hr_format) %>%
              mutate(SOCP2 = ifelse(SOCP2 == "Total","All Occupations", SOCP2)), 
            by = c("SOCP2_35hr"="SOCP2")) 


ggplot(df %>% filter(SEX_psrc=="Female"), aes(x=fct_relevel(reorder(SOCP2_35hr,PERNP_35hr_median),"All Occupations", after = Inf))) +
  geom_col(aes(y=share)) +
  geom_text(aes(y=share,label=share_format), 
            hjust = 0) +
  geom_text(aes(y=0.9,label=PERNP_35hr_format), 
            hjust = 0) +
  coord_flip() +
  scale_fill_manual(values = psrc_colors$pgnobgy_5)+
  scale_y_continuous(labels = scales::percent, limits = c(0,1.12)) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank())+
  labs(caption = pums_source)

ggplot(df %>% filter(SEX_psrc=="Female"), aes(x=fct_relevel(reorder(SOCP2_35hr,share), "All Occupations", after = Inf))) +
  geom_col(aes(y=share)) +
  geom_text(aes(y=share,label=share_format), 
            hjust = 0) +
  coord_flip() +
  scale_fill_manual(values = psrc_colors$pgnobgy_5)+
  scale_y_continuous(labels = scales::percent, limits = c(0,0.92)) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank(),
        axis.text.x = element_blank())+
  labs(caption = pums_source)
```

## data

```{r}
#| tbl-cap: share of women by occupations (workers working more than 35hr/wk) [2019-2023 ACS 5-year PUMS data]

df %>% 
  # mutate(share_pct = percent_format(share)) %>%
  arrange(desc(PERNP_35hr_median)) %>%
  pivot_wider(id_cols = c("SOCP2_35hr","PERNP_35hr_format"), names_from = "SEX_psrc", values_from = "share") %>%
  mutate(Female = percent_format(Female),
         Male = percent_format(Male))
```

:::

<!-- # HTS gender categories -->

<!-- ```{r} -->
<!-- #| tbl-cap: share of men and women [PSRC 2017-2023 Household Travel Survey] -->

<!-- psrc_hts_stat(df_hts,  -->
<!--               analysis_unit="person",  -->
<!--               group_vars=c("gender_final"), -->
<!--               incl_na = FALSE) %>% -->
<!--   mutate(prop_pct = percent_format(prop)) %>% -->
<!--   pivot_wider(id_cols = "gender_final", names_from = "survey_year", values_from = "prop_pct") -->
<!-- ``` -->


#	trip purpose by gender

- this chart shows the share of trips by trip purpose for female and male
- the results show that men makes more Work/School trips than women across all survey years, but the gap is narrowing over time
- in 2017/2019, women makes more Errand/Other, Escort trips than men
- in 2021, women makes more Shop trips than men

::: {.callout-note}
list of trip purposes that have statistically significant differences in share of trips between female and male:

- 2017/2019: 
  - f>m: Errand/Other, Escort
  - m>f: Work/School
- 2021:
  - f>m: Shop
  - m>f: Work/School
- 2023:
  - f>m: 
  - m>f: Work/School

:::

:::{.panel-tabset}

## chart

```{r}
#| fig-width: 7
#| fig-asp: 0.4
#| out-width: "90%"
#| fig-align: "center"

df <- psrc_hts_stat(df_hts, 
                     analysis_unit="trip", 
                     group_vars=c("gender_final", "purpose_custom"),
                     incl_na = FALSE) %>%
  mutate(prop_pct = percent_format(prop),
         prop_pct1 = percent_format(prop,1),
         purpose_custom_format = str_replace_all(purpose_custom,"/","\n"))

ggplot(df, aes(x=purpose_custom_format, y=prop, fill=gender_final)) +
  geom_col(position = position_dodge(), width=0.8) +
  facet_wrap(~ survey_year, ncol=4)+
  geom_text(aes(label=prop_pct1),
            vjust = -0.5,
            position = position_dodge(0.8)) +
  scale_fill_manual(values = psrc_colors$pgnobgy_5)+
  scale_y_continuous(labels = scales::percent, limits = c(0,0.37)) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank())+
  labs(caption = hts_source)


```

## chart with margin of error

```{r}
#| fig-width: 7
#| fig-asp: 0.6
#| out-width: "90%"
#| fig-align: "center"

ggplot(df, aes(x=purpose_custom_format, y=prop, fill=gender_final)) +
  geom_col(position = position_dodge(),width = 0.8) +
  geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                position = position_dodge(0.8),
                width = 0.4) +
  facet_wrap(~ survey_year, ncol=4)+
  scale_fill_manual(values = psrc_colors$pgnobgy_5)+
  scale_y_continuous(labels = scales::percent, limits = c(0,0.37)) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank())+
  labs(caption = hts_source)


```

## data

```{r}
#| tbl-cap: trip purpose by gender [Puget Sound Regional Council (PSRC) 2017-2023 Household Travel Survey]

df %>%
  pivot_wider(id_cols = c("survey_year","gender_final"), names_from = "purpose_custom", values_from = "prop_pct")
```

:::

# mode share by gender

- the chart shows the trip mode shares by gender
- the results show that women drive more than men in 2017/2019 and 2023
- men bike (in all years), walk (in 2017/2019 and 2023) and use transit (in 2023) more than women
- may pose the question: does this imply that women may have greater safety concerns when using these travel modes?

::: {.callout-note}
list of modes that have statistically significant differences in mode share between female and male:

- 2017/2019: 
  - f>m: Drive
  - m>f: Bike/Micromobility, Walk
- 2021:
  - f>m: 
  - m>f: Bike/Micromobility
- 2023:
  - f>m: Drive
  - m>f: Bike/Micromobility, Transit, Walk

:::

:::{.panel-tabset}

## chart

```{r}
#| fig-width: 7
#| fig-asp: 0.4
#| out-width: "90%"
#| fig-align: "center"

df <- psrc_hts_stat(df_hts, 
                     analysis_unit="trip", 
                     group_vars=c("gender_final", "mode_class_5"),
                     incl_na = FALSE) %>%
  mutate(prop_pct = percent_format(prop),
         prop_pct1 = percent_format(prop,1),
         mode_class_5_format = str_replace_all(mode_class_5,"/","\n") 
  )
test <- df %>% mutate(ymin = prop - prop_moe, ymax = prop + prop_moe)

ggplot(df, aes(x=mode_class_5_format, y=prop, fill=gender_final)) +
  geom_col(position = position_dodge(),width=0.8) +
  facet_wrap(~ survey_year, ncol=4)+
  geom_text(aes(label=prop_pct1),
            vjust = -0.5,
            position = position_dodge(0.8)) +
  scale_fill_manual(values = psrc_colors$pgnobgy_5)+
  scale_y_continuous(labels = scales::percent, limits = c(0,0.87)) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank())+
  labs(caption = hts_source)
```

## chart with margin of error

```{r}
#| fig-width: 7
#| fig-asp: 0.6
#| out-width: "90%"
#| fig-align: "center"

ggplot(df, aes(x=mode_class_5_format, y=prop, fill=gender_final)) +
  geom_col(position = position_dodge(),width = 0.8) +
  geom_errorbar(aes(ymin = prop - prop_moe, ymax = prop + prop_moe),
                position = position_dodge(0.8),
                width = 0.4) +
  facet_wrap(~ survey_year, ncol=4)+
  scale_fill_manual(values = psrc_colors$pgnobgy_5)+
  scale_y_continuous(labels = scales::percent, limits = c(0,0.87)) +
  psrc_style() +
  theme(panel.grid.major.y = element_blank(),
        axis.title = element_blank())+
  labs(caption = hts_source)


```

## data

```{r}
#| tbl-cap: mode share by gender [Puget Sound Regional Council (PSRC) 2017-2023 Household Travel Survey]

df %>%
  pivot_wider(id_cols = c("survey_year","gender_final"), names_from = "mode_class_5", values_from = "prop_pct")

```

:::

