---
title: "Hispanic Commute Modes"
output: html_document
fontsize: 12pt
---
\setmainfont{Poppins}
```{r setup, include=FALSE}
library(psrccensus)
library(ggplot2)
library(tidyverse)
library(psrcplot)
#source("C:/Users/cpeak/Repos/equity-focused-trends/hisp_heritage/2023/hisp_functions.R")
library(sf) # geographic data library
library(psrcelmer) # to get data from Elmer
library(psrcplot) # to get color schemes
library(mapview) # for outputting the map to html or png
library(htmlwidgets) # to help with map formatting
library(leafsync)

pums_2021 <- get_psrc_pums(span = 5,
                           dyear = 2021,
                           level = "p",
                           vars = c("AGEP","PRACE",
                                    "MIG",           # lived here 1 year ago
                                    "HISP",          # detailed Hispanic origin
                                    "HISPEED",       # Has broadband internet
                                    "FACCESSP",      # Access to the Internet
                                    "HHLDRHISP",     # Detailed Hispanic origin of householder
                                    "RAC2P",         # race with country
                                    "ANC1P",         # Ancestry - first entry
                                    "JWTRNS",        # means of transportation to work
                                    "JWMNP",         # Travel time to work
                                    "NAICSP",        # NAICS recode for 2018
                                    "POBP",          # Place of birth (Recode)
                                    "GRPIP",         # Gross rent as % of household income
                                    "OWN_RENT",      # Dichotomous tenure
                                    "MI_JOBSECTOR",  # PSRC-defined manufacturing-industrial groups
                                    "SOCP",          # Standard Occupational Classification (SOC) codes for 2018 and laterbased on 2018 SOC codes
                                    "TEN",           # Tenure
                                    "VEH",           # Vehicles (1 ton or less) available
                                    "WIF"))          # Workers in family

#pums_2021 <- pums_2021 %>%
#  hisp_pums_recode() %>%
#  mutate(ind = substr(NAICSP, 1,3))


#pums_race_counts <- pums_2021 %>%
#  psrc_pums_count(., group_vars="race_4cat") 

```

``` {r prep_map, echo=FALSE, message=FALSE, warning=FALSE}

create_psrc_map <- function(lyr,
                            lyr_data_field,
                            legend_title,
                            legend_subtitle,
                            psrc_col_pal,
                            map_lat = 47.615,
                            map_lon = -122.257,
                            map_zoom = 8.5,
                            wgs84 = 4326) {
  # psrc colors need more contrast to work
  pal <-
    leaflet::colorNumeric(palette = psrc_col_pal, domain = lyr_data_field)
  
  css_fix <-
    "div.info.legend.leaflet-control br {clear: both;} html * {font-family: Poppins !important;}" # CSS to correct spacing and font family
  html_fix <-
    htmltools::tags$style(type = "text/css", css_fix)  # Convert CSS to HTML
  
  
  labels <-
    paste0('Estimate: ', prettyNum(round(lyr_data_field,-1), big.mark = ",")) %>%
    lapply(htmltools::HTML)
  
  m <- leaflet::leaflet() %>%
    leaflet::addMapPane(name = "polygons", zIndex = 410) %>%
    leaflet::addMapPane(name = "maplabels", zIndex = 500) %>% # higher zIndex rendered on top
    
    leaflet::addProviderTiles("CartoDB.VoyagerNoLabels") %>%
    leaflet::addProviderTiles(
      "CartoDB.VoyagerOnlyLabels",
      options = leaflet::leafletOptions(pane = "maplabels"),
      group = "Labels"
    ) %>%
    leaflet::addEasyButton(leaflet::easyButton(
      icon = htmltools::span(class = "globe", htmltools::HTML("&#127758;")),  #&#127760; (another emoji option) #"fa-globe", (font awesome icon no longer works because of the conversion to Poppins font below)  
      title ="Region",
      onClick=JS("function(btn, map){map.setView([47.615,-122.257],8.5); }")))%>%
    leaflet::addPolygons(
      data = lyr,
      fillOpacity = 0.7,
      fillColor = pal(lyr_data_field),
      weight = 0.7,
      color = "#BCBEC0",
      group = "estimate",
      opacity = 0,
      stroke = FALSE,
      options = leaflet::leafletOptions(pane = "polygons"),
      dashArray = "",
      highlight = leaflet::highlightOptions(
        weight = 5,
        color = "76787A",
        dashArray = "",
        fillOpacity = 0.7,
        bringToFront = TRUE
      ),
      label = labels,
      labelOptions = leaflet::labelOptions(
        style = list(
          "font-weight" = "normal",
          padding = "3px 8px",
          "font-family" = "Poppins"
        )
      )
    ) %>%
    
    leaflet::addLegend(
      pal = pal,
      values = lyr_data_field,
      position = "bottomright",
      title = paste(legend_title, '<br>', legend_subtitle)
    ) %>%
    
    leaflet::addLayersControl(baseGroups = "CartoDB.VoyagerNoLabels",
                              overlayGroups = c("Labels", "estimate")) %>%
    
    leaflet::setView(lng = map_lon, lat = map_lat, zoom = map_zoom) %>%
    htmlwidgets::prependContent(html_fix)      # Insert into leaflet HTML code
  
  return(m)
  
}
big_tbl<-read.csv('tract_accessibility_2018.csv')
tbl <- big_tbl %>% mutate(geoid10=as.character(tract_geoid10))
tract_layer_name <- "TRACT2010_NOWATER"
lyr <- st_read_elmergeo(tract_layer_name)
lyr_data<- dplyr::left_join(lyr,tbl, by="geoid10")
# this is the field to map
lyr_data_field<-lyr_data$access
big_tbl <-psrccensus::get_acs_recs(geography='tract',table.names='B03001',year=2019, acs.type='acs5')
hisp_tbl <- big_tbl %>%filter(label=='Estimate!!Total:!!Hispanic or Latino:')%>% 
  dplyr::select(GEOID,estimate) %>%
  dplyr::mutate(dplyr::across(c('GEOID'), as.character))%>%
  dplyr::group_by(GEOID) %>%
  dplyr::summarise(Total=sum(estimate))
lyr_data_2<- dplyr::left_join(lyr_data,hisp_tbl, by = c("geoid10"='GEOID'))
# this is the field to map
lyr_data_field_2<-lyr_data_2$Total

```

### Intro 
	
Nationally, Hispanic populations [report relatively high use of public transit](https://tinyurl.com/5n83ta2s) compared with whites.  But in the central Puget Sound region, Census data shows that Hispanics tend to commute by car more often than other groups.  

```{r echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE}
pums_mode <- pums_2021 %>%
  filter(mode_hts != "NA") %>%
  psrc_pums_count(., group_vars=c("race_4cat","mode_hts")) %>%
  filter(mode_hts != "Total") 
pums_mode_drive <- pums_mode %>%
  filter(mode_hts == 'Drive')
  
mode_chart_drive <- interactive_column_chart(
  t=pums_mode_drive, y="share", x="mode_hts",
  fill="race_4cat",
  #facet="race_4cat",
  color='pognbgy_10',
  moe="share_moe",
  title="Drive Commute Mode choice by Race"
)
mode_chart_drive
```
Additionally, Hispanic transit commute rates tend to be roughly on par with whites and substantially below the rates for Black and Asian groups.  

```{r echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE}
pums_mode_transit <- pums_mode %>%
  filter(mode_hts == 'Transit')
mode_chart_transit <- interactive_column_chart(
  t=pums_mode_transit, y="share", x="mode_hts",
  fill="race_4cat",
  #facet="race_4cat",
  color='pognbgy_10',
  moe="share_moe",
  title="Transit Commute Mode choice by Race"
)
mode_chart_transit
```

This led us to wonder what are the factors that inform these choices, and what are some of the pressures that lead the area's Hispanics to favor cars over transit at higher rates than other populations.  

### Home Locations Make a Difference
First, we looked at the places that the Hispanic populace lives.  Using Census data from the American Community Survey, we looked at the Hispanic population counts by tract.  As shown in the map below, the roughly 441,000 Hispanic people in the region tend to live in either South King County -- notably Burien, Des Moines, Federal Way and Auburn -- and Southern Snohomish County along the SR 99 Corridor.  


```{r hispanic_pop_map, echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE}

psrc_purple_plus<-append(psrc_colors$purples_inc, "#FFFFFF", after=0)
map_acc<-create_psrc_map(lyr=lyr_data_2,
                        lyr_data_field=lyr_data_field,
                        legend_title= 'Transit Accessibility',
                        legend_subtitle = 'Number of Jobs within 45 minutes on Transit',
                        psrc_col_pal=psrc_purple_plus)

map_acc
# mapshot(map_acc, file='map_acc.png')


```

To see how well commuters in these heavily Hispanic areas are served by transit, we calculated the number of jobs accessible by transit within 45 minutes for each Census tract. This map, below, shows that the areas with best transit access to jobs are concentrated in central Seattle and Bellevue.  Residents in these areas can reach a large pool of jobs by a 45-minute transit ride -- nearly 800,000 for some tracts.  By contrast, people living in the most heavily Hispanic areas can reach a much smaller pool by equivalent means: many can reach only between 30 and 60 thousand within the same 45-minute window.  Residents of the tract with the single highest Hispanic population, in Auburn, have only 20,000 jobs accessible within this window  And the fewer the jobs accessible by transit, the higher the likelihood that a given worker will need to drive to get to his or her place of work.

```{r transit_access_map, echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE}
psrc_orange_plus<-append(psrc_colors$oranges_inc, "#FFFFFF", after=0)
 map_hisp_pop<-create_psrc_map(lyr=lyr_data_2, 
                         lyr_data_field=lyr_data_field_2,
                         legend_title= 'Hispanic Population',
                         legend_subtitle = 'ACS 2015-2019 5-year',
                         psrc_col_pal=psrc_orange_plus)

 map_hisp_pop
 # mapshot(map_hisp_pop, file='hisp_pop.png')
```

	
### Industry
We wondered if another contributing factor to the relatively high rates of Hispanic drive commute trips might be a difference in the mix of industrial sectors among the different groups.  To investigate this, we first summarized workers by the NAICS industry codes in Census PUMS data.  For Hispanic workers, we saw that construction jobs constitute the largest individual sector, with 7 percent of workers.  This was followed by Restaurants/Food Services (6.7 percent) and Elementary and Secondary Education (1.9 percent).  

For non-Hispanic workers, the Construction sector represented a lower share of the jobs, at 3.9 percent, although it was the largest single sector for this group just as it was for Hispanics.  Following this sector was Computer Systems Design (3.6 percent) and Restaurants/Food Services (3.5 percent).

```{r industry, echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE}
industry_shares_by_race <- pums_2021 %>%
  mutate(industry = factor(case_when(
          grepl('CON-Construction',NAICSP) ~ 'Construction',
          grepl('ENT-Restaurants', NAICSP) ~ 'Restaurants',
          grepl('PRF-Computer Systems Design', NAICSP) ~ 'Computer Systems Design',
          grepl('EDU-Elementary', NAICSP) ~ 'K-12 Education' )
        )
       ) %>%
  psrc_pums_count(., group_vars=c("race_4cat", "industry")) %>%
  filter(race_4cat != "Total") %>%
  filter(industry != "Total")

industry_shares_chart <- interactive_column_chart(
  t=industry_shares_by_race, y="share", x="race_4cat",
  fill="industry",
  pos="stack",
  color="pognbgy_5",
  #moe="share_moe",
  title = "Top Four Industries as Percent of Total Employment, by Race and Hispanic Origin"
)
industry_shares_chart

```
The relatively high rate of construction jobs among Hispanic workers is likely to have some marginal effect on the high drive commute share for this group.  As has been pointed out in the [news media](https://www.seattletimes.com/seattle-news/data/these-three-occupations-use-public-transit-far-more-than-most-in-king-county/), construction workers are among the least likely to commute by transit, not least because the work often require tools or equipment to be carried to the job site.  This is borne out in the Census PUMS data that we reviewed: only 3 percent of workers of in this sector (of any race) commute by transit, whereas 87 percent drive to work.  

It should be noted however that as relatively large as this employment share is, construction jobs still account for only seven percent of Hispanic employment, and are likely to be only a contributing factor to this group's high drive commute share.  And within the Construction sector itself, Hispanic commuters still showed a somewhat higher drive commute share than did other groups,with 91 percent of their commutes being by car.  The chart below illustrates this finding:

```{r commutes_by_race_cn, echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE}
pums_mode_cn <- pums_2021 %>%
  filter(mode_hts != "NA") %>%
  filter(grepl('Construction', NAICSP)) %>%
  psrc_pums_count(., group_vars=c("race_4cat","mode_hts")) %>%
  filter(mode_hts != "Total") 
  
mode_chart_cn_drive <- interactive_column_chart(
  t=pums_mode_cn %>% filter(mode_hts == 'Drive'), 
  y="share", x="mode_hts",
  fill="race_4cat",
  #facet="race_4cat",
  color='pognbgy_10',
  moe="share_moe",
  title="Drive Commute Mode choice by Race: Construction Only"
)
mode_chart_cn_drive
```

```{r scratch, echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE}
pums_naics_nonhisp<- pums_2021 %>%
  filter(race_4cat != "Hispanic or Latino") %>%
  psrc_pums_count(., group_vars=c("NAICSP")) %>%
  filter(NAICSP != "Total") %>%
  arrange(desc(count)) 

mode_by_naics <- pums_2021 %>%
  filter(!is.na(mode_hts)) %>%
  psrc_pums_count(., group_vars=c("NAICSP","mode_hts")) %>%
  filter(mode_hts != "Total") %>%
  filter(NAICSP != "Total")

# mode_by_ind2_rac4 <- mode_by_naics_rac4 

  
mode_by_ind2_rac4 <- pums_2021 %>%
  filter(!is.na(mode_hts)) %>%
  mutate(ind2 = factor(case_when(
          grepl('CON-Construction',NAICSP) ~ 'Construction',
          grepl('ENT-Restaurants', NAICSP) ~ 'Restaurants',
          grepl('PRF-Computer Systems Design', NAICSP) ~ 'Computer Systems',
          grepl('EDU-Elementary', NAICSP) ~ 'K-12 Education'#,
          #.default = ind )
          )
        )
       ) %>%
  psrc_pums_count(., group_vars=c("race_4cat", "ind2","mode_hts")) %>%
  filter(mode_hts != "Total") %>%
  filter(race_4cat != "Total") %>%
  filter(ind2 != "Total")

mode_by_race <- pums_2021 %>%
  filter(!is.na(mode_hts)) %>%
  mutate(ind2 = factor(case_when(
          grepl('CON-Construction',NAICSP) ~ 'Construction',
          grepl('ENT-Restaurants', NAICSP) ~ 'Restaurants',
          grepl('PRF-Computer Systems Design', NAICSP) ~ 'Computer Systems',
          grepl('EDU-Elementary', NAICSP) ~ 'K-12 Education'#,
          #.default = ind )
          )
        )
       ) %>%
  psrc_pums_count(., group_vars=c("race_4cat","mode_hts")) %>%
  filter(mode_hts != "Total") %>%
  filter(race_4cat != "Total")

mode_by_ind2_rac4_hisp <- mode_by_ind2_rac4 %>%
  filter(grepl('Hispanic', race_4cat))

mode_by_ind2_rac4_construction <- mode_by_ind2_rac4 %>%
  filter(ind2 == "Construction")

mode_by_naics_hisp <- pums_2021 %>%
  filter(!is.na(mode_hts)) %>%
  filter(race_4cat == "Hispanic or Latino") %>%
  psrc_pums_count(., group_vars=c("NAICSP","mode_hts")) %>%
  filter(mode_hts != "Total") %>%
  filter(NAICSP != "Total")

mode_by_naics_nonhisp <- pums_2021 %>%
  filter(!is.na(mode_hts)) %>%
  filter(race_4cat != "Hispanic or Latino") %>%
  psrc_pums_count(., group_vars=c("NAICSP","mode_hts")) %>%
  filter(mode_hts != "Total") %>%
  filter(NAICSP != "Total")


mode_splits_overall <- interactive_column_chart(
  t=mode_by_race,
  y="share", x="race_4cat",
  fill="mode_hts",
  pos="stack",
  color="pognbgy_5",
  title = paste("Commute Shares by Race")
)
#mode_splits_overall

mode_splits_chart <- function(industry) {
  mode_splits_for_industry <- interactive_column_chart(
    t=mode_by_ind2_rac4 %>% filter(ind2 == industry),
    y="share", x="race_4cat",
    fill="mode_hts",
    pos="stack",
    color="pognbgy_5",
    title = paste("Commute Shares by Race and Ethnicity:",industry)
  )
}
chart.construction <- mode_splits_chart('Construction')
#chart.construction



# chart.construction <- mode_splits_chart('Construction')
# chart.restaurants <- mode_splits_chart('Restaurants')
# chart.computers <- mode_splits_chart('Computer Systems')
# chart.med <- mode_splits_chart('MED')
# chart.construction
# chart.restaurants
# chart.computers
# chart.med



```


	


```{r echo = FALSE, cache=TRUE, warning=FALSE, message=FALSE}

pums_emp_mode <- pums_2021 %>%
  mutate(ind2 = factor(case_when(
          grepl('CON-Construction',NAICSP) ~ 'Construction',
          grepl('ENT-Restaurants', NAICSP) ~ 'Restaurants',
          #grepl('EDU-Elementary', NAICSP) ~ 'K-12 Edu',
          .default = 'Other' )
        )
       ) %>%
  filter(ind2 != 'Other') %>%
  psrc_pums_count(., group_vars=c("race_4cat", "ind2")) %>%
  inner_join(pums_race_counts, by='race_4cat') %>%
  mutate(pct = (count.x / count.y) * 100) %>%
  filter(ind2 != "Total")


industry_chart_all <- ggplot(
  pums_emp_mode, 
  aes(fill=ind2, y=pct, x=race_4cat)) +
    geom_bar(position="stack", stat="identity") 

#industry_chart_all
```