---
title: "Template to make PSRC leaflet map"
author: "PSRC Data"
date: "2023"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r, results='hide', message=FALSE, warning=FALSE}
library(sf) # geographic data library
library(dplyr) # for working with data frames
library(psrcelmer) # to get data from Elmer
library(psrcplot) # to get color schemes
library(mapview) # for outputting the map to html or png
library(htmlwidgets) # to help with map formatting
library(leafsync)
```


### 2. Read in any other libraries or keys you need

```{r, results='hide', message=FALSE, warning=FALSE}
library(psrccensus)
Sys.getenv("CENSUS_API_KEY")
```


```{r}
create_psrc_map <- function(lyr,
                            lyr_data_field,
                            legend_title,
                            legend_subtitle,
                            psrc_col_pal,
                            map_lat = 47.615,
                            map_lon = -122.257,
                            map_zoom = 8.5,
                            wgs84 = 4326) {
  # psrc colors need more contrast to work
  pal <-
    leaflet::colorNumeric(palette = psrc_col_pal, domain = lyr_data_field)
  
  css_fix <-
    "div.info.legend.leaflet-control br {clear: both;} html * {font-family: Poppins !important;}" # CSS to correct spacing and font family
  html_fix <-
    htmltools::tags$style(type = "text/css", css_fix)  # Convert CSS to HTML
  
  
  labels <-
    paste0('Estimate: ', prettyNum(round(lyr_data_field,-1), big.mark = ",")) %>%
    lapply(htmltools::HTML)
  
  m <- leaflet::leaflet() %>%
    leaflet::addMapPane(name = "polygons", zIndex = 410) %>%
    leaflet::addMapPane(name = "maplabels", zIndex = 500) %>% # higher zIndex rendered on top
    
    leaflet::addProviderTiles("CartoDB.VoyagerNoLabels") %>%
    leaflet::addProviderTiles(
      "CartoDB.VoyagerOnlyLabels",
      options = leaflet::leafletOptions(pane = "maplabels"),
      group = "Labels"
    ) %>%
    leaflet::addEasyButton(leaflet::easyButton(
      icon = htmltools::span(class = "globe", htmltools::HTML("&#127758;")),  #&#127760; (another emoji option) #"fa-globe", (font awesome icon no longer works because of the conversion to Poppins font below)  
      title ="Region",
      onClick=JS("function(btn, map){map.setView([47.615,-122.257],8.5); }")))%>%
    leaflet::addPolygons(
      data = lyr,
      fillOpacity = 0.7,
      fillColor = pal(lyr_data_field),
      weight = 0.7,
      color = "#BCBEC0",
      group = "estimate",
      opacity = 0,
      stroke = FALSE,
      options = leaflet::leafletOptions(pane = "polygons"),
      dashArray = "",
      highlight = leaflet::highlightOptions(
        weight = 5,
        color = "76787A",
        dashArray = "",
        fillOpacity = 0.7,
        bringToFront = TRUE
      ),
      label = labels,
      labelOptions = leaflet::labelOptions(
        style = list(
          "font-weight" = "normal",
          padding = "3px 8px",
          "font-family" = "Poppins"
        )
      )
    ) %>%
    
    leaflet::addLegend(
      pal = pal,
      values = lyr_data_field,
      position = "bottomright",
      title = paste(legend_title, '<br>', legend_subtitle)
    ) %>%
    
    leaflet::addLayersControl(baseGroups = "CartoDB.VoyagerNoLabels",
                              overlayGroups = c("Labels", "estimate")) %>%
    
    leaflet::setView(lng = map_lon, lat = map_lat, zoom = map_zoom) %>%
    htmlwidgets::prependContent(html_fix)      # Insert into leaflet HTML code
  
  return(m)
  
}

```


### 3. Define create_psrc_map function


### 4.  Read in Data Table with a geographic field ID

```{r, results=FALSE, message=FALSE, warning=FALSE}
big_tbl<-read.csv('C:/GitHub/equity-focused-trends/hisp_heritage/2023/tract_accessibility_2018.csv')
```

### 5. Clean up the data in preparation to join to the geographic layer
```{r}
# read in data table and process data to join and map
tbl <- big_tbl %>% mutate(geoid10=as.character(tract_geoid10))
```


### 6. Read in ElmerGeo geographic layer

```{r}
tract_layer_name <- "TRACT2010_NOWATER"
lyr <- st_read_elmergeo(tract_layer_name)
```

### 7. Join to data table layer via matching IDs, choose the field to map
```{r }
lyr_data<- dplyr::left_join(lyr,tbl, by="geoid10")
# this is the field to map
lyr_data_field<-lyr_data$access
```


```{r, results=FALSE, message=FALSE, warning=FALSE}
big_tbl <-psrccensus::get_acs_recs(geography='tract',table.names='B03001',year=2019, acs.type='acs5')
```

### 5. Clean up the data in preparation to join to the geographic layer
```{r}
# read in data table and process data to join and map
hisp_tbl <- big_tbl %>%filter(label=='Estimate!!Total:!!Hispanic or Latino:')%>% 
  dplyr::select(GEOID,estimate) %>%
  dplyr::mutate(dplyr::across(c('GEOID'), as.character))%>%
  dplyr::group_by(GEOID) %>%
  dplyr::summarise(Total=sum(estimate))
```

```{r }
lyr_data_2<- dplyr::left_join(lyr_data,hisp_tbl, by = c("geoid10"='GEOID'))
# this is the field to map
lyr_data_field_2<-lyr_data_2$Total
```

### 8. Call the function to make the map.

these are the function inputs:

lyr: sf map layer, the geographic layer including the field you want to map

lyr_data_field: numeric vector of data values to map

legend_title: string, The title in the legend

legend_subtitle: string, The title in the legend

psrc_col_pal: string list, color ramp

map_lat=47.615, map_lon=-122.257: defaults, can be overridden for centering the map

map_zoom=8.5, wgs84=4326 : defaults, can be overriden, zoom level and projection
```{r}
# need to add a white color to ensure contrast
psrc_purple_plus<-append(psrc_colors$purples_inc, "#FFFFFF", after=0)
map_acc<-create_psrc_map(lyr=lyr_data_2, 
                        lyr_data_field=lyr_data_field,
                        legend_title= 'Transit Accessibility',
                        legend_subtitle = 'Number of Jobs within 45 minutes on Transit',
                        psrc_col_pal=psrc_purple_plus)

map_acc
mapshot(map_acc, file='map_acc.png')
```

```{r}
# need to add a white color to ensure contrast
psrc_orange_plus<-append(psrc_colors$oranges_inc, "#FFFFFF", after=0)
map_hisp_pop<-create_psrc_map(lyr=lyr_data_2, 
                        lyr_data_field=lyr_data_field_2,
                        legend_title= 'Hispanic Population',
                        legend_subtitle = 'ACS 2015-2019 5-year',
                        psrc_col_pal=psrc_orange_plus)

map_hisp_pop
mapshot(map_hisp_pop, file='hisp_pop.png')
```

```{r}
map_it<-sync(map_acc, map_hisp_pop, ncol=1)
map_it
```






