---
title: "R Notebook"
output: html_notebook
---

map for population density
[x] transit mode share by county
commute mode share by county (only aapi)
aggregate PI in place of birth charts


```{r}
library(psrc.travelsurvey)
library(psrccensus)
library(psrcplot)
library(tidycensus)
library(psrcelmer)
library(psrcctpp)

library(tidyverse)
library(gridExtra)
library(sf)

install_psrc_fonts()

options(scipen=999)
# vars_meta$variable

pums_caption <- "Source: U.S. Census Bureau, American Community Survey (ACS) 2019 and 2021 1-Year Public Use\n Microdata Sample (PUMS) Data in King, Kitsap, Pierce and Snohomish counties"


# create workbook for data download


```

```{r}
tract.asian <-get_acs_recs(geography='tract',table.names=c('B02011'),years=c(2021))
tract.na.pi <-get_acs_recs(geography='tract',table.names=c('B02012'),years=c(2021))

tract.lyr <- st_read_elmergeo('TRACT2020_NOWATER')

create_tract_map(tract.tbl=tract.asian, tract.lyr=tract.lyr,
map.title.position='topleft', legend.title='Asian Alone or in Combination Population',
legend.subtitle='ACS 2021 (5 yr)')
```


```{r}
create_tract_map(tract.tbl=tract.na.pi, tract.lyr=tract.lyr,
map.title.position='topleft', legend.title='Native Hawaiian and other Pacific Islander alone or in Combination Population',
legend.subtitle='ACS 2021 (5 yr)')
```

# transit and walking frequency
```{r}

trans <- hhts_count(per_data_17_19, 
                   group_vars=c("race_aapi", "transit_freq2"),
                   spec_wgt = "hh_weight_2017_2019_adult",
           incl_na = FALSE) %>%
  add_row(hhts_count(per_data_21, 
                   group_vars=c("race_aapi", "transit_freq2"),
                   spec_wgt = "person_adult_weight_2021",
           incl_na = FALSE)) %>%
  filter(transit_freq2=="at least 1 day/week",
         #race_aapi %in% c("Asian or Pacific Islander","White only")
         )

tele <- hhts_count(per_data_17_19,
           group_vars=c("race_aapi", "workplace_travel"),
           spec_wgt = "hh_weight_2017_2019_adult",
           incl_na = FALSE) %>%
  add_row(
    hhts_count(per_data_21,
               group_vars=c("race_aapi", "workplace_travel"),
               spec_wgt = "person_adult_weight_2021",
               incl_na = FALSE)) %>%
  filter(#race_aapi %in% c("Asian or Pacific Islander","White only"),
         workplace_travel=="Works at home")

ggplot(trans, aes(x=race_aapi, y=share, fill=survey)) +
  geom_col(position = "dodge")+  
  # facet_wrap(~race_aapi)+
  moe_bars+
  scale_y_continuous(labels=percent) +
  scale_fill_discrete_psrc ("gnbopgy_5")+
  labs(title="Transit usage at least 1 day/week by race", caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
  psrc_style() + 
  theme(axis.title.x= element_blank(),
          axis.title.y= element_blank(),
          axis.text.x = element_text(angle=15,vjust=0.9,hjust=0.6))


ggplot(tele, aes(x=race_aapi, y=share, fill=survey)) +
  geom_col(position = "dodge")+
  # facet_wrap(~race_aapi)+
  moe_bars+
  scale_y_continuous(labels=percent) +
  scale_fill_discrete_psrc ("gnbopgy_5")+
  labs(title="Telework by race", caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
          axis.title.y= element_blank(),
          axis.text.x = element_text(angle=15,vjust=0.9,hjust=0.6))



```



# trip mode share by race
```{r}
ggplot(hts_mode_by_race_4cat, 
       aes(x= mode_simple, y= share, fill= survey)) +
  geom_col(position = "dodge") +
  facet_wrap(~race_4cat)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$pognbgy_10)+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Trip Mode Share", caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))
ggplot(hts_mode_by_race_3cat, 
       aes(x= mode_simple, y= share, fill= survey)) +
  geom_col(position = "dodge") +
  facet_wrap(~race_3cat)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$pognbgy_10)+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Trip Mode Share", caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))

```

# trip mode share by race (each mode)
```{r}
all_modes <- list("Drive","Transit","Walk","Bike")

# commute mode share (no WFH) by race (4 categories)
plot_trip_mode_share_by_race <- function(plot_mode){
  
    ggplot(hts_mode_by_race_4cat %>% filter(mode_simple == plot_mode), 
           aes(x= race_4cat, y= share, fill= survey)) +
      geom_col(position = "dodge") +
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$purples_inc[c(2,4)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
      labs(title=paste("Trip Mode Share (", plot_mode, ") by race", sep=""),
      caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank(),
            axis.text.x = element_text(angle=15, hjust=0.7))
}
for (mode in all_modes){print(plot_trip_mode_share_by_race(mode))}

# commute mode share (no WFH) by race (3 categories)
plot_trip_mode_share_by_race <- function(plot_mode){
  
    ggplot(hts_mode_by_race_3cat %>% filter(mode_simple == plot_mode), 
           aes(x= race_3cat, y= share, fill= survey)) +
      geom_col(position = "dodge") +
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$blues_inc[c(3,5)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
      labs(title=paste("Trip Mode Share (", plot_mode, ") by race", sep=""),
      caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank())
}
for (mode in all_modes){print(plot_trip_mode_share_by_race(mode))}

```

# show only 2021
```{r}

print_plot<-ggplot(hts_mode_by_race_3cat %>% filter(mode_simple %in% c("Drive","Transit","Walk"),
                                        survey == "2021") %>%
       mutate(race_3cat = case_when(
         race_3cat=="Asian or Pacific Islander"~"Asian or \nPacific \nIslander",
         race_3cat=="Other people of color"~"Other \npeople of \ncolor",
         TRUE~race_3cat)), 
       aes(x= race_3cat, y= share, fill = race_3cat)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$pognbgy_10)+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  facet_wrap(~mode_simple, scales = "free_y")+
  labs(title=paste("Trip Mode Share by Race", sep=""),
  caption = "2021 PSRC Household Travel Survey")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())

```



# mode share of each trip purpose by race
```{r}
plot_trip_mode_share_by_race <- function(plot_mode, plot_purpose){
ggplot(hts_mode_purpose_by_race_3cat %>% 
         filter(simple_purpose==plot_purpose,mode_simple == plot_mode), 
           aes(x= race_3cat, y= share, fill= survey)) +
      geom_col(position = "dodge") +
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$blues_inc[c(3,5)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
      labs(title=paste(plot_purpose, " Trip Mode Share (", plot_mode, ") by Race", sep=""),
      caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank())
  
}

for (mode in unique(hts_mode_purpose_by_race_3cat$mode_simple[c(1,2,4)])){
  for (purpose in unique(hts_mode_purpose_by_race_3cat$simple_purpose)){
    
  print(plot_trip_mode_share_by_race(mode,purpose))
  }
  }
```

```{r}

for (purpose in unique(hts_mode_purpose_by_race_3cat$simple_purpose)){
print(
  ggplot(hts_mode_purpose_by_race_3cat %>%
           filter(simple_purpose==purpose,
                  !mode_simple %in% c("Bike","Other")) %>%
           mutate(race_3cat = case_when(
             race_3cat=="Asian or Pacific Islander"~"Asian or \nPacific \nIslander",
             race_3cat=="Other people of color"~"Other \npeople of \ncolor",
             TRUE~race_3cat)), 
           aes(x= race_3cat, y= share, fill= survey)) +
      geom_col(position = "dodge") +
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$blues_inc[c(3,5)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
  facet_wrap(~mode_simple, scales = "free_y")+
      labs(title=paste(purpose, " Trip Mode Share by Race", sep=""),
      caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank()))
}

for (purpose in unique(hts_mode_purpose_by_race_3cat2$purpose_work_other)){
print(
  ggplot(hts_mode_purpose_by_race_3cat2 %>%
           filter(purpose_work_other==purpose,
                  !mode_simple %in% c("Bike","Other")) %>%
           mutate(race_3cat = case_when(
             race_3cat=="Asian or Pacific Islander"~"Asian or \nPacific \nIslander",
             race_3cat=="Other people of color"~"Other \npeople of \ncolor",
             TRUE~race_3cat)), 
           aes(x= race_3cat, y= share, fill= survey)) +
      geom_col(position = "dodge") +
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$purples_inc[c(3,5)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
  facet_wrap(~mode_simple, scales = "free_y")+
      labs(title=paste(purpose, " Trip Mode Share by Race", sep=""),
      caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank()))
}

# purpose="Shop"
# purpose="Errands"
# purpose="Social/Recreation/Meal"
# print_plot<-ggplot(hts_mode_purpose_by_race_3cat %>%
#            filter(simple_purpose==purpose,
#                   !mode_simple %in% c("Bike","Other")) %>%
#            mutate(race_3cat = case_when(
#              race_3cat=="Asian or Pacific Islander"~"Asian or \nPacific \nIslander",
#              race_3cat=="Other people of color"~"Other \npeople of \ncolor",
#              TRUE~race_3cat)), 
#            aes(x= race_3cat, y= share, fill= survey)) +
#       geom_col(position = "dodge") +
#       scale_y_continuous(labels = scales::label_percent()) +
#       scale_fill_manual(values = psrc_colors$blues_inc[c(3,5)])+
#       geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
#                               width=0.2, position = position_dodge(0.9))+
#   facet_wrap(~mode_simple, scales = "free_y")+
#       labs(title=paste(purpose, " Trip Mode Share by Race", sep=""),
#       caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
#       psrc_style() +
#       theme(axis.title.x= element_blank(),
#             axis.title.y= element_blank())
```




# commute mode share by race
```{r}
ggplot(pums_race_commute_19_4cat %>%
         add_row(pums_race_commute_21_4cat) %>% 
         filter(!mode %in% c("Total",NA,"Other"),
                race_4cat %in% c("Asian or Pacific Islander","White alone")), 
       aes(x= mode, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  facet_wrap(~race_4cat)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$pognbgy_10)+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Commute Mode Share", caption = "2019 and 2021 Public Use Microsample Census data")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))

ggplot(pums_race_commute_19_no_wfm_4cat %>%
         add_row(pums_race_commute_21_no_wfm_4cat) %>% 
         filter(!mode %in% c("Total","Drive",NA,"Other"),
                race_4cat %in% c("Asian or Pacific Islander","White alone")), 
       aes(x= mode, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  facet_wrap(~race_4cat)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$pognbgy_10)+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Commute Mode Share", caption = "2019 and 2021 Public Use Microsample Census data")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))
```

# commute mode share by race [each mode]


```{r}
library(openxlsx)
wb <- createWorkbook()
xlsx_file <- "data_spreadsheet.xlsx"
plot <- pums_race_commute_19_4cat %>%
         add_row(pums_race_commute_21_4cat) %>% 
         filter(mode %in% c("Worked from home")) %>%
         mutate(race_4cat = factor(case_when(race_4cat == "Black or African American alone"~"Black or African \nAmerican",
                                             race_4cat == "White alone"~"White",
                                             race_4cat == "Hispanic or Latino"~"Hispanic or \nLatinx",
                                             race_4cat == "Asian or Pacific Islander"~"Asian or Pacific \nIslander",
                                             race_4cat == "Some Other Race(s)"~"Some Other \nRace(s)",
                                             TRUE~race_4cat), 
                                   levels = c("Asian or Pacific \nIslander",
                                              "Black or African \nAmerican",
                                              "Hispanic or \nLatinx",
                                              "White",
                                              "Some Other \nRace(s)")))

test2 <- pums_race_commute_19_3cat_county %>%
           add_row(pums_race_commute_21_3cat_county) %>% 
           filter(mode %in% c("Worked from home"),
                  race_3cat == "Asian or Pacific Islander") %>%
           add_row(pums_race_commute_19_3cat %>%
                     add_row(pums_race_commute_21_3cat) %>% 
                     filter(mode %in% c("Worked from home"),
                            race_3cat == "Asian or Pacific Islander")) %>%
           mutate(COUNTY = factor(COUNTY,
                                  levels=c("Region","King","Kitsap","Pierce","Snohomish")))

test3 <- pums_country_commute_21 %>%
         filter(mode %in% c("Worked from home"),
                !country %in% c("All combinations of Asian races only","Other Asian alone","Total"),
                share_moe>0,
                count>1500) %>%
         mutate(country = gsub(" alone", "", country),
                country = case_when(country == "Chinese, except Taiwanese,"~ "Chinese",
                                     TRUE~country),
                country = str_wrap(country, width=23),
                country = fct_reorder(country, share))

addWorksheet(wb = wb, sheetName = 'Telework by Race and Ethnicity', gridLines = FALSE)
writeDataTable(wb = wb, sheet = 'Telework by Race and Ethnicity', x = plot)
addWorksheet(wb = wb, sheetName = 'Telework by County', gridLines = FALSE)
writeDataTable(wb = wb, sheet = 'Telework by County', x = test2)
addWorksheet(wb = wb, sheetName = 'Telework by AAPI Groups', gridLines = FALSE)
writeDataTable(wb = wb, sheet = 'Telework by AAPI Groups', x = test3)


saveWorkbook(wb, xlsx_file, overwrite = TRUE)
```


```{r}
# share of WFH by race (AAPI and other races)
ggplot(plot, 
       aes(x= race_recat, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent(), limits=c(0,0.4)) +
  scale_fill_manual(values = psrc_colors$blues_inc[c(3,5)])+
  # geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
  #                         width=0.2, position = position_dodge(0.9))+
  labs(title="Share of Telework by Race And Ethnicity", caption = pums_caption)+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())

# share of WFH by race (AAPI, other POC and white)
ggplot(pums_race_commute_19_3cat  %>%
         add_row(pums_race_commute_21_3cat ) %>% 
         filter(mode %in% c("Worked from home")), 
       aes(x= race_3cat, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$oranges_inc[c(2,4)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Share of Telework by Race", 
       caption = "2019 and 2021 American Community Survey (ACS) Public Use Microdata Sample (PUMS) data")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())

all_modes <- list("Drive","Public Transit","Walked","Bicycle")

# commute mode share (no WFH) by race (4 categories)
plot_commute_mode_share_by_race <- function(plot_mode){
  
  return(
    ggplot(pums_race_commute_19_no_wfm_4cat  %>%
           add_row(pums_race_commute_21_no_wfm_4cat ) %>% 
           filter(mode %in% c(plot_mode)), 
           aes(x= race_4cat, y= share, fill= factor(DATA_YEAR))) +
      geom_col(position = "dodge") +
      # facet_wrap(~mode,ncol = 4,scales = "free_y")+
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$purples_inc[c(2,4)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
      labs(title=paste("Commute Mode Share (", plot_mode, ") by Race", sep=""), 
           caption = "2019 and 2021 American Community Survey (ACS) Public Use Microdata Sample (PUMS) data")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank(),
              axis.text.x = element_text(angle=15, hjust=0.7))
    )

}
for (mode in all_modes){print(plot_commute_mode_share_by_race(mode))}


# commute mode share (no WFH) by race (3 categories)
plot_commute_mode_share_by_race <- function(plot_mode){
  
  return(
    ggplot(pums_race_commute_19_no_wfm_3cat %>%
           add_row(pums_race_commute_21_no_wfm_3cat) %>% 
           filter(mode %in% c(plot_mode)), 
           aes(x= race_3cat, y= share, fill= factor(DATA_YEAR))) +
      geom_col(position = "dodge") +
      # facet_wrap(~mode,ncol = 4,scales = "free_y")+
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$blues_inc[c(3,5)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
      labs(title=paste("Commute Mode Share (", plot_mode, ") by Race", sep=""), 
           caption = "2019 and 2021 American Community Survey (ACS) Public Use Microdata Sample (PUMS) data")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank())
    )

}
for (mode in all_modes){print(plot_commute_mode_share_by_race(mode))}

```
```{r}
pdf(file="Share of AAPI Teleworkers by County.pdf",height=4.32,width=7)
print(
  ggplot(pums_race_commute_19_3cat_county %>%
           add_row(pums_race_commute_21_3cat_county) %>% 
           filter(mode %in% c("Worked from home"),
                  race_3cat == "Asian or Pacific Islander") %>%
           add_row(pums_race_commute_19_3cat %>%
                     add_row(pums_race_commute_21_3cat) %>% 
                     filter(mode %in% c("Worked from home"),
                            race_3cat == "Asian or Pacific Islander")) %>%
           mutate(COUNTY = factor(COUNTY,
                                  levels=c("Region","King","Kitsap","Pierce","Snohomish"))),
         aes(x= COUNTY, y= share, fill= factor(DATA_YEAR))) +
    geom_col(position = "dodge") +
    # geom_text(
    #   aes(label = scales::percent(share, accuracy = 1), y = share + 0.005),
    #   position = position_dodge(0.9),
    #   vjust = 0
    # ) +
    # facet_wrap(~COUNTY,ncol = 4)+
    scale_y_continuous(labels = scales::label_percent()) +
    scale_fill_manual(values = psrc_colors$purples_inc[c(2,4)])+
    # geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
    #                         width=0.2, position = position_dodge(0.9))+
    labs(title="Share of AAPI Teleworkers by County", caption = pums_caption)+
    psrc_style() +
    theme(axis.title.x= element_blank(),
          axis.title.y= element_blank())
)
dev.off()
```
# telework share by race by county
```{r}

# share of WFH by race (AAPI and other races)
ggplot(pums_race_commute_19_3cat_county %>%
         add_row(pums_race_commute_21_3cat_county) %>% 
         filter(mode %in% c("Worked from home"),
                race_3cat == "Asian or Pacific Islander") %>%
         add_row(pums_race_commute_19_3cat %>%
                   add_row(pums_race_commute_21_3cat) %>% 
                   filter(mode %in% c("Worked from home"),
                          race_3cat == "Asian or Pacific Islander")) %>%
         mutate(COUNTY = factor(COUNTY,
                                levels=c("Region","King","Kitsap","Pierce","Snohomish"))),
       aes(x= COUNTY, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  # geom_text(
  #   aes(label = scales::percent(share, accuracy = 1), y = share + 0.005),
  #   position = position_dodge(0.9),
  #   vjust = 0
  # ) +
  # facet_wrap(~COUNTY,ncol = 4)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(2,4)])+
  # geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
  #                         width=0.2, position = position_dodge(0.9))+
  labs(title="Share of Telework in AAPI by County", caption = pums_caption)+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())

plot_commute_mode_share_by_race_by_county <- function(plot_mode){
  
  return(
    ggplot(pums_race_commute_19_no_wfm_3cat_county %>%
             add_row(pums_race_commute_21_no_wfm_3cat_county) %>% 
             filter(mode %in% c(plot_mode)), 
             aes(x= race_3cat, y= share, fill= factor(DATA_YEAR))) +
      geom_col(position = "dodge") +
      facet_wrap(~COUNTY,ncol = 4)+
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$blues_inc[c(3,5)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
      labs(title=paste("Commute Mode Share (",plot_mode,") by race", sep=""), 
           caption = "2019 and 2021 Public Use Microsample Census data")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank(),
              axis.text.x = element_text(angle=15,vjust=1.1, hjust=1.05))
  )

}

all_modes <- list("Drive","Public Transit","Bicycle")

for (mode in all_modes){print(plot_commute_mode_share_by_race_by_county(mode))}
```

```{r}
ggplot(pums_race_commute_19_no_wfm_3cat_county %>%
             add_row(pums_race_commute_21_no_wfm_3cat_county) %>% 
             filter(mode == "Public Transit",
                    race_3cat == "Asian or Pacific Islander"), 
             aes(x= COUNTY, y= share, fill= factor(DATA_YEAR))) +
      geom_col(position = "dodge") +
      # facet_wrap(~COUNTY,ncol = 4)+
      scale_y_continuous(labels = scales::label_percent()) +
      scale_fill_manual(values = psrc_colors$blues_inc[c(3,5)])+
      geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                              width=0.2, position = position_dodge(0.9))+
      labs(title="Transit Commute Mode Share in AAPI by County", 
           caption = "2019 and 2021 Public Use Microsample Census data")+
      psrc_style() +
      theme(axis.title.x= element_blank(),
            axis.title.y= element_blank(),
              axis.text.x = element_text(angle=15,vjust=1.1, hjust=1.05))
```


# demographic characteristics
```{r car ownership in HTS and PUMS}

# share of WFH by race (AAPI and other races)
ggplot(pums_race_veh_21 %>%
         filter(vehicle == "No vehicle"), 
       aes(x= race_3cat, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(4)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Car-free household by race", caption = "2021 Public Use Microsample Census data")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())
ggplot(pums_race_veh_21_per %>%
         filter(vehicle == "No vehicle"), 
       aes(x= race_3cat, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(4)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Car-free person by race", caption = "2021 Public Use Microsample Census data")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())


# ggplot(pums_race_veh_21_county %>%
#          filter(vehicle == "No vehicle"), 
#        aes(x= race_3cat, y= share, fill= factor(DATA_YEAR))) +
#   geom_col(position = "dodge") +
#   facet_wrap(~COUNTY,ncol = 4)+
#   scale_y_continuous(labels = scales::label_percent()) +
#   scale_fill_manual(values = psrc_colors$greens_inc[c(4)])+
#   geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
#                           width=0.2, position = position_dodge(0.9))+
#   labs(title="Car-free household by race by county", caption = "2021 Public Use Microsample Census data")+
#   psrc_style() +
#   theme(axis.title.x= element_blank(),
#         axis.title.y= element_blank(),
#           axis.text.x = element_text(angle=15,vjust=1.1, hjust=1))

ggplot(hts_car_own %>%
         filter(vehicle_count == "No vehicle") %>%
         mutate(survey = factor(survey, levels = c("2017","2019","2021"))), 
       aes(x= race_3cat, y= share, fill= survey)) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(1,3,5)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Car-free person by race in 2017, 2019 and 2021 [person-level]", caption = "2017, 2019 and 2021 PSRC Household Travel Survey")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())

ggplot(hts_car_own %>%
         filter(vehicle_count == "No vehicle",
                survey == "2021"), 
       aes(x= race_3cat, y= share, fill= survey)) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(3,5)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Car-free person by race [person-level]", caption = "2017, 2019 and 2021 PSRC Household Travel Survey")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())

plot <- hts_car_own %>% 
  filter(vehicle_count == "No vehicle",
         survey == "2021") %>%
  select(race_3cat,share,share_moe) %>%
  mutate(data = "2021 PSRC HTS") %>%
  add_row(pums_race_veh_21_per %>%
            filter(vehicle == "No vehicle") %>%
            select(race_3cat,share,share_moe) %>%
            mutate(data = "2021 1-Year PUMS"))

ggplot(plot, 
       aes(x= race_3cat, y= share, fill= data)) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(2,5)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Car-free person by race [person-level]", caption = "2021 Public Use Microsample Census data and 2021 PSRC Household Travel Survey")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())
```
```{r}
ggplot(hts_commute %>%
         filter(commute_mode3!="Total",commute_mode3!="Other") %>%
         mutate(survey = factor(survey, levels = c("2017","2019","2021"))), 
       aes(x= race_3cat, y= share, fill= survey)) +
  geom_col(position = "dodge") +
  facet_wrap(~commute_mode3,scales = "free_y")+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(1,3,5)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Commute mode by race in 2017, 2019 and 2021 [person-level]", caption = "2017, 2019 and 2021 PSRC Household Travel Survey")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank())
ggplot(hts_commute2 %>%
         filter(commute_mode4!="Total",commute_mode4!="Other") %>%
         mutate(survey = factor(survey, levels = c("2017","2019","2021"))), 
       aes(x= race_3cat, y= share, fill= survey)) +
  geom_col(position = "dodge") +
  facet_wrap(~commute_mode4,scales = "free_y")+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(1,3,5)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Commute mode by race in 2017, 2019 and 2021 [person-level]", caption = "2017, 2019 and 2021 PSRC Household Travel Survey")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))

plot<-pums_race_commute_17_no_wfm_3cat %>%
         add_row(pums_race_commute_19_no_wfm_3cat) %>%
         add_row(pums_race_commute_21_no_wfm_3cat) %>%
         mutate(DATA_YEAR = factor(DATA_YEAR, levels = c("2017","2019","2021"))) %>%
  filter(race_3cat!="Total",mode_hts!="Total",mode_hts!="Other")
ggplot(plot, 
       aes(x= race_3cat, y= share, fill= DATA_YEAR)) +
  geom_col(position = "dodge") +
  facet_wrap(~mode_hts,scales = "free_y")+
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(1,3,5)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Commute mode by race in 2017, 2019 and 2021 [person-level]", caption = "2017, 2019 and 2021 Public Use Microsample Census data")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))
```
```{r}
commute_race_compare <- hts_commute_race_compare %>%
  add_row(pums_commute_compare) %>%
  filter(mode_hts!="Total")%>%
  mutate(race_3cat = case_when(
                      race_3cat=="Asian or Pacific Islander"~"Asian or \nPacific \nIslander",
                      race_3cat=="Other people of color"~"Other \npeople of \ncolor",
                      race_3cat=="White alone"~"White \nalone"))

for (mode in unique(commute_race_compare$mode_hts)){
  print(
    ggplot(commute_race_compare %>% filter(mode_hts == mode), 
         aes(x= race_3cat, y= share, fill= data_type)) +
    geom_col(position = "dodge") +
    facet_wrap(~DATA_YEAR,scales = "free_y")+
    scale_y_continuous(labels = scales::label_percent()) +
    scale_fill_manual(values = psrc_colors$greens_inc[c(3,5)])+
    geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                            width=0.2, position = position_dodge(0.9))+
    labs(title=paste(mode, "commute mode by race in 2017, 2019 and 2021 "), caption = "2017, 2019 and 2021 Public Use Microsample Census data and PSRC Household Travel Survey")+
    psrc_style() +
    theme(axis.title.x= element_blank(),
          axis.title.y= element_blank()))
}
```
```{r}
mode = "Drive"
mode = "Transit"
mode = "Walk/Bike"
mode = "Other"
print_plot <-ggplot(commute_race_compare %>% filter(mode_hts == mode), 
         aes(x= race_3cat, y= share, fill= data_type)) +
    geom_col(position = "dodge") +
    facet_wrap(~DATA_YEAR,scales = "free_y")+
    scale_y_continuous(labels = scales::label_percent()) +
    scale_fill_manual(values = psrc_colors$greens_inc[c(3,5)])+
    geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                            width=0.2, position = position_dodge(0.9))+
    labs(title=paste(mode, "commute mode by race in 2017, 2019 and 2021 "), caption = "2017, 2019 and 2021 Public Use Microsample Census data and PSRC Household Travel Survey")+
    psrc_style() +
    theme(axis.title.x= element_blank(),
          axis.title.y= element_blank())
print_plot
```


```{r}

print_plot <- ggplot(pums_race_age_21 %>% 
         filter(age!="Total") %>%
         mutate(age = factor(age, levels = c("Under 18 years","18-24 years",
                                             "25-34 years","35-44 years",
                                             "45-54 years","55-64 years",
                                             "65 years and above"   ))), 
       aes(x=age, y= share, fill=race_3cat)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::label_percent()) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(1,3,5)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="Age by Race", caption = "2021 Public Use Microsample Census data")+
  psrc_style() +
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
        axis.text.x = element_text(angle=15,vjust=1.1, hjust=0.8))

```


```{r}
jpeg(file="Share of Teleworkers for 10 Most Populous AAPI Groups.jpeg",height=4.32,width=7)
print(
  ggplot(pums_country_commute_21 %>%
         filter(mode %in% c("Worked from home"),
                !country %in% c("All combinations of Asian races only","Other Asian alone","Total"),
                share_moe>0,
                count>1500) %>%
         mutate(country2 = gsub(" alone", "", country),
                country2 = case_when(country2 == "Chinese, except Taiwanese,"~ "Chinese",
                                     TRUE~country2),
                country2 = str_wrap(country2, width=23),
                country2 = fct_reorder(country2, share)), 
       aes(x= country2, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent(),limits = c(0,NA)) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(4)])+
  # geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
  #                         width=0.2, position = position_dodge(0.9))+
  labs(title="Share of Teleworkers for 10 Most Populous AAPI Groups", caption = "Source: U.S. Census Bureau, American Community Survey (ACS) 2021 1-Year Public Use Microdata\n Sample (PUMS) Data in King, Kitsap, Pierce and Snohomish counties", x="")+
  psrc_style()+
    coord_flip()+
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color="#cbcbcb"),
      legend.position = "None"
    )
)
dev.off()
```

```{r}

# share of WFH by race (AAPI and other races)
ggplot(pums_race_migrate_21 %>%
         filter(mode %in% c("Worked from home"),
                share_moe>0) %>%
         mutate(migrate = fct_reorder(migrate, share)), 
       aes(x= migrate, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent(),limits = c(0,NA)) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(4)])+
  geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
                          width=0.2, position = position_dodge(0.9))+
  labs(title="[AAPI] Share of Telework by Place of Birth", caption = "2021 Public Use Microsample Census data", x="")+
  psrc_style()+
    coord_flip()+
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color="#cbcbcb")
    )


  
ggplot(pums_country_commute_21 %>%
         filter(mode %in% c("Worked from home"),
                !country %in% c("All combinations of Asian races only","Other Asian alone","Total"),
                share_moe>0,
                count>1500) %>%
         mutate(country2 = gsub(" alone", "", country),
                country2 = case_when(country2 == "Chinese, except Taiwanese,"~ "Chinese",
                                     TRUE~country2),
                country2 = str_wrap(country2, width=23),
                country2 = fct_reorder(country2, share)), 
       aes(x= country2, y= share, fill= factor(DATA_YEAR))) +
  geom_col(position = "dodge") +
  # facet_wrap(~race_aapi)+
  scale_y_continuous(labels = scales::label_percent(),limits = c(0,NA)) +
  scale_fill_manual(values = psrc_colors$greens_inc[c(4)])+
  # geom_errorbar(aes(ymin=share-share_moe, ymax=share+share_moe),
  #                         width=0.2, position = position_dodge(0.9))+
  labs(title="Share of Telework in 10 Largest AAPI Groups by Population", caption = "Source: U.S. Census Bureau, American Community Survey (ACS) 2021 1-Year Public Use Microdata\n Sample (PUMS) Data in King, Kitsap, Pierce and Snohomish Counties", x="")+
  psrc_style()+
    coord_flip()+
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color="#cbcbcb"),
      legend.position = "None"
    )

```


```{r}

ggplot(purpose_plot %>% filter(race_aapi %in% c("Asian or Pacific Islander","White only")), 
            aes(x=simple_purpose, y=trips_per_person, fill=survey)) +
  geom_col(position = "dodge")+  
  facet_wrap(~race_aapi)+
  geom_errorbar(aes(ymin=trips_per_person-moe_trips_person, ymax=trips_per_person+moe_trips_person),
                width=0.2, position = position_dodge(0.9))+
  scale_y_continuous(limits = c(0,2))+
  scale_fill_discrete_psrc ("gnbopgy_5")+
  labs(title="Trip Purpose", caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
  psrc_style()+
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))

ggplot(purpose_plot %>% filter(!race_aapi %in% c("Asian or Pacific Islander","White only")), 
            aes(x=simple_purpose, y=trips_per_person, fill=survey)) +
  geom_col(position = "dodge")+  
  facet_wrap(~race_aapi)+
  geom_errorbar(aes(ymin=trips_per_person-moe_trips_person, ymax=trips_per_person+moe_trips_person),
                width=0.2, position = position_dodge(0.9))+
  scale_y_continuous(limits = c(0,2))+
  scale_fill_discrete_psrc ("gnbopgy_5")+
  labs(title="Trip Purpose", caption = "2017/2019 and 2021 PSRC Household Travel Surveys")+
  psrc_style()+
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))
```

```{r}
ggplot(acs_race_occupation_5year %>% 
         filter(name=="Region",
                occupation!="Total" ) %>%
         wrap_axis(occupation,20) %>%
         mutate(race = stringr::str_wrap(race, width=15)), 
       aes(x= fct_rev(cat), y= occupation_share, fill= race)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = label_percent()) +
  scale_fill_manual(values = psrc_colors$pognbgy_10)+
  geom_errorbar(aes(ymin=occupation_share-occupation_moe, ymax=occupation_share+occupation_moe),
                          width=0.2, position = position_dodge(0.9))+
  coord_flip()+
  psrc_style() +
  labs(title="Share of Workers by Industry", caption = "2021 Public Use Microsample Census data")+
  theme(axis.title.x= element_blank(),
        axis.title.y= element_blank(),
        panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color="#cbcbcb"),
      legend.position = "right")

```


```{r}

```


```{r difference between 1 and 5 year data}
diff <- race_commute_5year %>%
  ungroup() %>%
  add_row(race_commute_1year %>%
  ungroup()) %>%
  filter(race %in% c("Asian only","Native Hawaiian and other Pacific Islander alone"),
         name == "Region",
         mode!="Total")%>%
  mutate(mode = factor(mode, levels=c("Drove alone",
                                       "Carpooled",
                                       "Public transit",
                                       "Walked",
                                       "Other",
                                       "Work from home")),
         acs_type = factor(acs_type,levels=c("acs5","acs1")))

ggplot(diff, aes(x=mode,y=estimate,fill=acs_type))+
  geom_col(position="dodge")+
  scale_fill_manual(values = psrc_colors$pognbgy_10)+
  geom_errorbar(aes(ymin=estimate-moe, ymax=estimate+moe),
                            width=0.2, position = position_dodge(0.9))+
    facet_wrap(~race, scales = "free_y")+
  psrc_style()+
  theme(axis.title.x= element_blank(),
          axis.title.y= element_blank(),
          axis.text.x = element_text(angle=45,vjust=1.1, hjust=1.05))
```

# Place of Birth by race
```{r}
test <- get_acs_recs(geography = "county", table.names = 'B06004I', years = 2021, acs.type = 'acs5')
  
plot_acs_birth <- function(.data, acs_type){
  .plot_data <- .data %>% 
   mutate(born = stringr::str_wrap(born, width=11),
          race = stringr::str_wrap(race, width=22),
          born = factor(born, levels = c("Born in WA\nstate",
                                         "Born in\nother state",
                                         "Native:\nborn\noutside the\nUS",
                                         "Foreign\nborn")),
          race = factor(race,levels = c("American Indian and\nAlaska Native alone",
                                        "Asian only",
                                        "Black or African\nAmerican alone",
                                        "Hispanic or Latino",
                                        "Native Hawaiian and\nother Pacific Islander\nalone",
                                        "White alone")),
          born_moe = ifelse(is.na(born_moe),0,born_moe))
  ggplot(.plot_data %>% filter(name=="Region"), 
       aes(x= born, y= born_share, fill= race)) +
    geom_col(position = "dodge") +
    scale_y_continuous(labels = label_percent()) +
    scale_fill_manual(values = psrc_colors$pognbgy_10)+
    geom_errorbar(aes(ymin=born_share-born_moe, ymax=born_share+born_moe),
                            width=0.2, position = position_dodge(0.9))+
    psrc_style() +
    labs(title = paste("Place of birth share by race - ACS 2021 [", acs_type, "year] data"))+
    theme(axis.title.x= element_blank(),
          axis.title.y= element_blank())
}
plot_acs_birth(race_birthplace_5year, "5")
plot_acs_birth(race_birthplace_1year, "1")
```
#old analysis
```{r}
plot <- data_commute %>% 
   filter(!commute_mode2 %in% c("NA", "Total", "Other modes")) %>%
   wrap_axis(commute_mode2, w=15) %>%
   mutate(cat = factor(cat, levels = c("Drive alone","HOV modes",
                "Public transit","Walk","Bike or micro-\nmobility")))


ggplot(plot, aes(x= cat, y= share, fill= race_aapi)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = label_percent()) +
  scale_fill_manual(values = psrc_colors$pognbgy_10)+
  moe_bars+
  psrc_style() +
  labs(title = "Commute mode share by race")+
  theme(axis.title.x= element_blank())
```

```{r ACS commute by race}
# TODO: change to facet like diff charts
plot_acs_commute <- function(.data, acs_type){
  .plot_data <- .data %>% 
   filter(!mode %in% c("Total")) %>%
   mutate(race = stringr::str_wrap(race, width=22),
          race = factor(race,levels = c("American Indian and\nAlaska Native alone",
                                        "Asian only",
                                        "Black or African\nAmerican alone",
                                        "Hispanic or Latino",
                                        "Native Hawaiian and\nother Pacific Islander\nalone",
                                        "White alone")),
          mode = factor(mode, levels=c("Drove alone",
                                       "Carpooled",
                                       "Public transit",
                                       "Walked",
                                       "Other",
                                       "Work from home")))

  ggplot(.plot_data %>% filter(name=="Region"), aes(x= mode, y= mode_share, fill= race)) +
    geom_col(position = "dodge") +
    scale_y_continuous(labels = label_percent()) +
    scale_fill_manual(values = psrc_colors$pognbgy_10)+
    geom_errorbar(aes(ymin=mode_share-mode_moe, ymax=mode_share+mode_moe),
                            width=0.2, position = position_dodge(0.9))+
    psrc_style() +
    labs(title = paste("Commute mode share by race - ACS 2021 [", acs_type, "year] data"))+
    theme(axis.title.x= element_blank(),
          axis.title.y= element_blank())
}
plot_acs_commute(race_commute_5year, "5")
plot_acs_commute(race_commute_1year, "1")
```